### 2020-04-28 paper.tex
Do not add project name into file names

\def\paperversiondraft{draft}
\def\paperversionblind{blind}
\def\paperversioncamera{camera}

% If no special paper-version is requested, compile in draft mode
\ifx\paperversion\paperversionblind
\else
  \ifx\paperversion\paperversioncamera
  \else
     \def\paperversion{draft}
  \fi
\fi

\def\grammarlyon{on}

\ifx\grammarly\grammarlyon
\def\review{}
\else
\def\review{review,}
\fi

\ifx\paperversion\paperversiondraft
  \documentclass[sigplan,\review anonymous]{acmart}
\fi

\ifx\paperversion\paperversionblind
  \documentclass[sigplan,\review anonymous]{acmart}
\fi

\ifx\paperversion\paperversioncamera
  \documentclass[sigplan]{acmart}\settopmatter{}
\fi

\usepackage{xargs}
\usepackage{todonotes}
\usepackage{xparse}
\usepackage{xifthen, xstring}
\usepackage{ulem}
\usepackage{xspace}
\makeatletter
\font\uwavefont=lasyb10 scaled 652

\newcommand\colorwave[1][blue]{\bgroup\markoverwith{\lower3\p@\hbox{\uwavefont\textcolor{#1}{\char58}}}\ULon}
\makeatother

\ifx\paperversion\paperversiondraft
\newcommand\createtodoauthor[2]{%
\def\tmpdefault{emptystring}
\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
   {\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}}
   {\ifthenelse{\equal{##2}{}}{\colorwave[#2]{##1}\xspace}{ \todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}\colorwave[#2]{##1}}}}}
\else
\newcommand\createtodoauthor[2]{%
\expandafter\newcommand\csname #1\endcsname[2][\@nil]{}}
\fi

% Broaden margins to make room for todo notes
\ifx\paperversion\paperversiondraft
  \paperwidth=\dimexpr \paperwidth + 6cm\relax
  \oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
  \evensidemargin=\dimexpr\evensidemargin + 3cm\relax
  \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
  \setlength{\marginparwidth}{4cm}
\fi

% We use the following color scheme
% 
% This scheme is both print-friendly and colorblind safe for
% up to four colors (including the red tones makes it not
% colorblind safe any more)
%
% https://colorbrewer2.org/#type=qualitative&scheme=Paired&n=4

\definecolor{pairedOneLightBlue}{HTML}{a6cee3}
\definecolor{pairedTwoDarkBlue}{HTML}{1f78b4}
\definecolor{pairedThreeLightGreen}{HTML}{b2df8a}
\definecolor{pairedFourDarkGreen}{HTML}{33a02c}
\definecolor{pairedFiveLightRed}{HTML}{fb9a99}
\definecolor{pairedSixDarkRed}{HTML}{e31a1c}

\createtodoauthor{grosser}{pairedOneLightBlue}
\createtodoauthor{authorTwo}{pairedTwoDarkBlue}
\createtodoauthor{authorThree}{pairedThreeLightGreen}
\createtodoauthor{authorFour}{pairedFourDarkGreen}
\createtodoauthor{authorFive}{pairedFiveLightRed}
\createtodoauthor{authorSix}{pairedSixDarkRed}

%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format should change 'sigplan' to
%% 'acmsmall'.

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

\begin{document}

%% Title information
\title[Short Title]{Full Title}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Subtitle}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Paper note
%% The \thanks command may be used to create a "paper note" ---
%% similar to a title note or an author note, but not explicitly
%% associated with a particular element.  It will appear immediately
%% above the permission/copyright statement.
\thanks{with paper note}                %% \thanks is optional
                                        %% can be repeated if necesary
                                        %% contents suppressed with 'anonymous'

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

% Only add ACM notes and keywords in camera ready version
% Drop citations and footnotes in draft and blind mode.
\ifx\paperversion\paperversioncamera
%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional

\else
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\fi

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle
\ifx\grammarly\grammarlyon 
\onecolumn 
\else 
\fi

\section{Introduction}

Of this super cool article \ldots
\grosser{A simple comment refering to a particular location}

If there is \authorTwo[some text we want to refer to specifically]{this looks
particularly interesting} we can refer to this text from within our comments.

And here we only underline \authorTwo[Text]{} that is interesting, but do not
actually comment.

Now we have a sequence of comments,
\authorThree{Comment 3}
\authorFour{Comment 4}
\authorFive{Comment 5}
\authorSix{Comment 6}
which should not introduce overlong white space sequences.

~\\
Our contributions are:
\grosser{Always state your contributions explicitly: (A) this makes it
easy for the reader to understand what novelty is presented, and (B)
these contributions help you to focus. In particular, the objective of
the remaining paper should be to support the claims stated here.
}
\begin{itemize}
	\item Contribution 1
	\item Contribution 2
	\item Contribution 3
	\item Contribution 4
\end{itemize}

\section{Related Work}

\grosser{Related work should always be at the end of the document,
         as it otherwise becomes an obstacle your reader must
	 overcome before reaching your idea. For details see:
	 \url{https://www.microsoft.com/en-us/research/academic-program/write-great-research-paper/}
}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
%\bibliography{bibfile}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

### 2020-04-28 initialize.py

@@ -1,44 +1,0 @@
-#!/usr/bin/env python
-
-import os
-
-directory = os.path.dirname(os.path.realpath(__file__))
-directory = os.path.basename(directory)
-
-projectname = directory
-
-os.rename("paper-template.tex", projectname + ".tex")
-
-
-def replaceFile(filename):
-    f = open(filename,'r')
-    filedata = f.read()
-    f.close()
-
-    newdata = filedata.replace("paper-template", projectname)
-
-    f = open(filename + ".bak",'w')
-    f.write(newdata)
-    f.close()
-    os.rename(filename + ".bak", filename)
-
-replaceFile("Makefile")
-replaceFile("README.md")
-replaceFile(".github/workflows/build_paper.yml")
-replaceFile("paper-template-grammarly.tex")
-replaceFile("paper-template-blind.tex")
-replaceFile("paper-template-camera.tex")
-
-os.rename("paper-template-grammarly.tex", projectname + "-grammarly.tex")
-os.rename("paper-template-blind.tex", projectname + "-blind.tex")
-os.rename("paper-template-camera.tex", projectname + "-camera.tex")
-
-from subprocess import call
-
-call(['git', 'add', '*.tex'])
-call(['git', 'add', '*.pdf'])
-call(['git', 'rm', 'initialize.py'])
-call(['git', 'add', 'README.md'])
-call(['git', 'add', 'Makefile'])
-call(['git', 'add', '.github'])
-call(['git', 'commit', '-m', 'Initialize project'])

### 2020-10-22 create-source-archive.sh
deploy: 74a1e82fcdf9662a70667365dd8b31320abbc653

@@ -1,0 +1,2 @@
+#!/usr/bin/env bash
+git archive  --prefix="paper-source/" HEAD . > paper-source.zip

### 2020-10-22 create-source-archive.sh
Actually use 'zip' as output format

@@ -1,2 +1,2 @@
 #!/usr/bin/env bash
-git archive  --prefix="paper-source/" HEAD . > paper-source.zip
+git archive  --prefix="paper-source/" --format=zip HEAD . > paper-source.zip

### 2020-11-22 tools/create-video.py
Build video from pdfs when new release is tagged

@@ -1,0 +1,100 @@
+#!/usr/bin/env python3
+
+from subprocess import run
+import os
+import math
+import sys
+import shutil
+
+def getPageNumber(name):
+    files = os.listdir(os.path.dirname(name))
+    files = filter(lambda f: f.startswith(os.path.basename(name)), files)
+    files = filter(lambda f: f.endswith('.png'), files)
+    files = list(files)
+    return len(files)
+
+def removePNGFiles(directory):
+    files = os.listdir(directory)
+    files = filter(lambda x: x.find(".png") != -1, files)
+    for f in files:
+      os.remove(directory + '/' + f)
+
+def createImageOfPaper(path, width = 6, pages = 18):
+    print(path)
+    removePNGFiles(os.path.dirname(path))
+
+    name = path
+
+    cmd = ['convert', '-density', '300', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + ".png"]
+    run(cmd)
+
+    pageNumber = getPageNumber(path);
+    if pageNumber == 1:
+        shutil.copyfile(name + ".png", name + "-0.png")
+
+    for i in range(pageNumber, pages):
+        cmd = ['convert', '-size', '1020x1320', 'xc:white', name + '-' + str(i) + '.png']
+        run(cmd)
+
+    for row in range(0, math.ceil(pages / width)):
+        cmd = ['convert']
+        for imageId in range(row * width, min((row + 1) * width, pages)):
+            cmd.append(name + "-" + str(imageId) + ".png")
+        cmd.append("+append")
+        cmd.append(name + "-row-" + str(row) + ".png")
+        run(cmd)
+
+    cmd = ['convert']
+    for row in range(0, math.ceil(pages / width)):
+      cmd.append(name + "-row-" + str(row) + ".png")
+    cmd.append("-append")
+    cmd.append(name + "-full.png")
+    run(cmd)
+
+def getReleases(path):
+    releases = os.listdir(path)
+    releases = filter(lambda x: x.find('master-release') != -1, releases)
+    releases = map(lambda x: x.replace('master-release-', ''), releases)
+    releases = map(lambda x: int(x), releases)
+    releases = list(releases)
+    releases.sort()
+    releases = map(lambda x: path + 'master-release-' + str(x) + "/blind.pdf", releases)
+    releases = list(releases)
+    releases = filter(os.path.exists, releases)
[…]

### 2020-11-22 tools/create-video.py
deploy: 6ce3fe3a9111a072f02a47f4c8d3c4e4ff23d437

@@ -84,7 +84,7 @@ def moveImages(path):
 
 def createVideo(path):
     run(['rm', path + 'video.mp4'])
-    cmd = ['ffmpeg', '-r', '1', '-i', path + 'still-%05d.png', '-vcodec', 'libx264', '-crf', '25', '-pix_fmt', 'yuv420p', path + 'video.mp4']
+    cmd = ['ffmpeg', '-r', '1', '-i', path + 'still-%05d.png', '-vcodec', 'libx264', '-crf', '25', '-pix_fmt', 'yuv420p', '-vf', 'scale=3840:2160', path + 'video.mp4']
     run(cmd)
     print(path + 'video.mp4')
 

### 2020-11-22 tools/create-video.py
deploy: 4e6e540dea22e029e035b75decaffe0958372215

@@ -19,7 +19,7 @@ def removePNGFiles(directory):
     for f in files:
       os.remove(directory + '/' + f)
 
-def createImageOfPaper(path, width = 6, pages = 18):
+def createImageOfPaper(path, width = 7, pages = 21):
     print(path)
     removePNGFiles(os.path.dirname(path))
 

### 2020-11-22 tools/create-video.py
deploy: d10ae1951a6ce640b7def74a4878784e0b09a6ac

@@ -13,9 +13,12 @@ def getPageNumber(name):
     files = list(files)
     return len(files)
 
-def removePNGFiles(directory):
+def removePNGFiles(directory, tmpOnly = False):
     files = os.listdir(directory)
     files = filter(lambda x: x.find(".png") != -1, files)
+    if tmpOnly:
+        files = filter(lambda x: x.find("-tmp") != -1, files)
+
     for f in files:
       os.remove(directory + '/' + f)
 
@@ -25,12 +28,12 @@ def createImageOfPaper(path, width = 7, pages = 21):
 
     name = path
 
-    cmd = ['convert', '-density', '300', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + ".png"]
+    cmd = ['convert', '-density', '300', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
     run(cmd)
 
     pageNumber = getPageNumber(path);
     if pageNumber == 1:
-        shutil.copyfile(name + ".png", name + "-0.png")
+        shutil.copyfile(name + ".png", name + "-tmp-0.png")
 
     for i in range(pageNumber, pages):
         cmd = ['convert', '-size', '1020x1320', 'xc:white', name + '-' + str(i) + '.png']
@@ -41,14 +44,15 @@ def createImageOfPaper(path, width = 7, pages = 21):
         for imageId in range(row * width, min((row + 1) * width, pages)):
             cmd.append(name + "-" + str(imageId) + ".png")
         cmd.append("+append")
-        cmd.append(name + "-row-" + str(row) + ".png")
+        cmd.append(name + "-tmp-row-" + str(row) + ".png")
         run(cmd)
 
     cmd = ['convert']
     for row in range(0, math.ceil(pages / width)):
-      cmd.append(name + "-row-" + str(row) + ".png")
+      cmd.append(name + "-tmp-row-" + str(row) + ".png")
     cmd.append("-append")
     cmd.append(name + "-full.png")
+    removePNGFiles(path, tmpOnly=True)
     run(cmd)
 
 def getReleases(path):

### 2020-11-22 tools/create-video.py
Use directory name

@@ -52,7 +52,7 @@ def createImageOfPaper(path, width = 7, pages = 21):
       cmd.append(name + "-tmp-row-" + str(row) + ".png")
     cmd.append("-append")
     cmd.append(name + "-full.png")
-    removePNGFiles(path, tmpOnly=True)
+    removePNGFiles(os.path.dirname(path), tmpOnly=True)
     run(cmd)
 
 def getReleases(path):

### 2020-11-22 tools/create-video.py
Delete temporary files properly

@@ -36,13 +36,13 @@ def createImageOfPaper(path, width = 7, pages = 21):
         shutil.copyfile(name + ".png", name + "-tmp-0.png")
 
     for i in range(pageNumber, pages):
-        cmd = ['convert', '-size', '1020x1320', 'xc:white', name + '-' + str(i) + '.png']
+        cmd = ['convert', '-size', '1020x1320', 'xc:white', name + '-tmp-' + str(i) + '.png']
         run(cmd)
 
     for row in range(0, math.ceil(pages / width)):
         cmd = ['convert']
         for imageId in range(row * width, min((row + 1) * width, pages)):
-            cmd.append(name + "-" + str(imageId) + ".png")
+            cmd.append(name + "-tmp-" + str(imageId) + ".png")
         cmd.append("+append")
         cmd.append(name + "-tmp-row-" + str(row) + ".png")
         run(cmd)
@@ -52,8 +52,8 @@ def createImageOfPaper(path, width = 7, pages = 21):
       cmd.append(name + "-tmp-row-" + str(row) + ".png")
     cmd.append("-append")
     cmd.append(name + "-full.png")
-    removePNGFiles(os.path.dirname(path), tmpOnly=True)
     run(cmd)
+    removePNGFiles(os.path.dirname(path), tmpOnly=True)
 
 def getReleases(path):
     releases = os.listdir(path)

### 2020-11-22 tools/create-video.py
Revert "Video: Create PDFs in parallel"
This reverts commit 5a18499acbba80b486d96a844a8e077002492e87.

@@ -68,13 +68,12 @@ def getReleases(path):
     releases = list(releases)
     return list(releases)
 
-from multiprocessing import Pool
 
 def createImages(path):
     releases = getReleases(path)
 
-    with Pool(4) as p:
-        p.map(createImageOfPaper, releases)
+    for release in releases:
+        createImageOfPaper(release)
 
 def moveImages(path):
     removePNGFiles(path)

### 2020-11-22 tools/create-video.py
deploy: eaeddf36aa2dcfbbde830e3c449079101f03f9d7

@@ -32,6 +32,7 @@ def createImageOfPaper(path, width = 7, pages = 21):
     run(cmd)
 
     pageNumber = getPageNumber(path);
+    run(['ls', os.path.dirname(path)])
     if pageNumber == 1:
         shutil.copyfile(name + ".png", name + "-tmp-0.png")
 

### 2020-11-22 tools/create-video.py
deploy: 61ddae08323f6af90cce0e8b3714603921040310

@@ -32,9 +32,8 @@ def createImageOfPaper(path, width = 7, pages = 21):
     run(cmd)
 
     pageNumber = getPageNumber(path);
-    run(['ls', os.path.dirname(path)])
     if pageNumber == 1:
-        shutil.copyfile(name + ".png", name + "-tmp-0.png")
+        shutil.copyfile(name + "-tmp.png", name + "-tmp-0.png")
 
     for i in range(pageNumber, pages):
         cmd = ['convert', '-size', '1020x1320', 'xc:white', name + '-tmp-' + str(i) + '.png']

### 2020-11-22 tools/create-video.py
deploy: 8c5b50fdedca1acc79a30d8aa9484595d1a2f8f4

@@ -68,12 +68,13 @@ def getReleases(path):
     releases = list(releases)
     return list(releases)
 
+from multiprocessing import Pool
 
 def createImages(path):
     releases = getReleases(path)
 
-    for release in releases:
-        createImageOfPaper(release)
+    with Pool(4) as p:
+        p.map(createImageOfPaper, releases)
 
 def moveImages(path):
     removePNGFiles(path)

### 2020-11-30 plots/speedup.py
deploy: 214686dfc36e2038e661ead28dce9c57168c1999

@@ -17,12 +17,22 @@ labels = ['G1', 'G2', 'G3', 'G4', 'G5']
 men_means = [1.5, 1.2, 1.3, 1.1, 1.0]
 women_means = [1.8, 1.5, 1.1, 1.3, 0.9]
 
+# Color palette
+light_gray = "#cacaca"
+dark_gray = "#827b7b"
+light_blue = "#a6cee3"
+dark_blue = "#1f78b4"
+light_green = "#b2df8a"
+dark_green = "#33a02c"
+light_red = "#fb9a99"
+dark_red = "#e31a1c"
+
 x = np.arange(len(labels))  # the label locations
 width = 0.35  # the width of the bars
 
 fig, ax = plt.subplots()
-rects1 = ax.bar(x - width/2, men_means, width, label='Men', color = "#a6cee3")
-rects2 = ax.bar(x + width/2, women_means, width, label='Women', color = "#1f78b4")
+rects1 = ax.bar(x - width/2, men_means, width, label='Men', color = light_blue)
+rects2 = ax.bar(x + width/2, women_means, width, label='Women', color = dark_blue)
 
 # Y-Axis Label
 #

### 2021-04-21 paper.tex
Publications accepted

@@
 \ifx\paperversion\paperversiondraft
-  \documentclass[sigplan,\review anonymous]{acmart}
+  \documentclass[manuscript,\review anonymous]{acmart}
 \fi
@@
 \ifx\paperversion\paperversionblind
-  \documentclass[sigplan,\review anonymous]{acmart}
+  \documentclass[manuscript,\review anonymous]{acmart}
 \fi
@@
 \ifx\paperversion\paperversioncamera
-  \documentclass[sigplan]{acmart}\settopmatter{}
+  \documentclass[manuscript, nonacm]{acmart}\settopmatter{}
 \fi
 
+\usepackage{colortbl}
+
+% 'draftonly' environment
+\usepackage{environ}
+\ifx\paperversion\paperversiondraft
+\newenvironment{draftonly}{}{}
+\else
+\NewEnviron{draftonly}{}
+\fi
+
+% Most PL conferences are edited by conference-publishing.com. Follow their
+% advice to add the following packages.
+%
+% The first enables the use of UTF-8 as character encoding, which is the
+% standard nowadays. The second ensures the use of font encodings that support
+% accented characters etc. (Why should I use this?). The mictotype package
+% enables certain features 'to­wards ty­po­graph­i­cal per­fec­tion
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{microtype}
+
 \usepackage{xargs}
-\usepackage{todonotes}
+\usepackage{lipsum}
+\usepackage[textsize=tiny]{todonotes}
 \usepackage{xparse}
 \usepackage{xifthen, xstring}
-\usepackage{ulem}
+\usepackage[normalem]{ulem}
 \usepackage{xspace}
+\usepackage{marginnote}
 \makeatletter
@@
    {\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}}
-   {\ifthenelse{\equal{##2}{}}{\colorwave[#2]{##1}\xspace}{ \todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}\colorwave[#2]{##1}}}}}
+   {\ifthenelse{\equal{##2}{}}{\colorwave[#2]{##1}\xspace}{\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}\colorwave[#2]{##1}}}}
+\expandafter\newcommand\csname #1f\endcsname[2][\tmpdefault]{
+	\marginnote{
+		\todo[inline,linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1 (Figure):} ##2}}
+   }
+}
+
 \else
 \newcommand\createtodoauthor[2]{%
-\expandafter\newcommand\csname #1\endcsname[2][\@nil]{}}
+\def\tmpdefault{emptystring}
+\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
+   {}
+   {\ifthenelse{\equal{##2}{}}{##1\xspace}{##1}}}
+\expandafter\newcommand\csname #1f\endcsname[2][\tmpdefault]{
+   }
+   }
 \fi
@@
 % Broaden margins to make room for todo notes
+\makeatletter
+\patchcmd{\@addmarginpar}{\ifodd\c@page}{\ifodd\c@page\@tempcnta\m@ne}{}{}
+\makeatother
 \ifx\paperversion\paperversiondraft
+  \makeatletter
+  \if@ACM@journal
+    \geometry{asymmetric}
+    \paperwidth=\dimexpr \paperwidth + 3.5cm\relax
+    \oddsidemargin=\dimexpr\oddsidemargin + 0cm\relax
+    \evensidemargin=\dimexpr\evensidemargin + 0cm\relax
+    \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
+    \setlength{\marginparwidth}{4.6cm}
+    % This makeatletter box helps to move notes to the right
+    \makeatletter
+    \long\def\@mn@@@marginnote[#1]#2[#3]{%
+      \begingroup
+        \ifmmode\mn@strut\let\@tempa\mn@vadjust\else
+          \if@inlabel\leavevmode\fi
+          \ifhmode\mn@strut\let\@tempa\mn@vadjust\else\let\@tempa\mn@vlap\fi
+        \fi
+        \@tempa{%
+          \vbox to\z@{%
+            \vss
+            \@mn@margintest
+            \if@reversemargin\if@tempswa
+                \@tempswafalse
+              \else
+                \@tempswatrue
+            \fi\fi
+            %\if@tempswa
+              \rlap{%
+                \if@mn@verbose
+                  \PackageInfo{marginnote}{xpos seems to be \@mn@currxpos}%
+                \fi
+                \begingroup
+                  \ifx\@mn@currxpos\relax\else\ifx\@mn@currxpos\@empty\else
+                      \kern-\dimexpr\@mn@currxpos\relax
+                  \fi\fi
+                  \ifx\@mn@currpage\relax
+                    \let\@mn@currpage\@ne
+                  \fi
+                  \if@twoside\ifodd\@mn@currpage\relax
+                      \kern\oddsidemargin
+                    \else
+                      \kern\evensidemargin
+                    \fi
+                  \else
+                    \kern\oddsidemargin
+                  \fi
+                  \kern 1in
+                \endgroup
+                \kern\marginnotetextwidth\kern\marginparsep
+                \vbox to\z@{\kern\marginnotevadjust\kern #3
+                  \vbox to\z@{%
+                    \hsize\marginparwidth
+                    \linewidth\hsize
+                    \kern-\parskip
+                    \marginfont\raggedrightmarginnote\strut\hspace{\z@}%
+                    \ignorespaces#2\endgraf
+                    \vss}%
+                  \vss}%
+              }%
+          }%
+        }%
+      \endgroup
+    }
+    \makeatother
+  \else
     \paperwidth=\dimexpr \paperwidth + 6cm\relax
@@
     \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
-  \setlength{\marginparwidth}{4cm}
+    \setlength{\marginparwidth}{4.6cm}
+  \fi
+  \makeatother
 \fi
@@
 
+\definecolor{pairedNegOneLightGray}{HTML}{cacaca}
+\definecolor{pairedNegTwoDarkGray}{HTML}{827b7b}
 \definecolor{pairedOneLightBlue}{HTML}{a6cee3}
@@
 \createtodoauthor{authorSix}{pairedSixDarkRed}
+
+\graphicspath{{./images/}}
+
+% Define macros that are used in this paper
+%
+% We require all macros to end with a delimiter (by default {}) to enusure
+% that LaTeX adds whitespace correctly.
+\makeatletter
+\newcommand\requiredelimiter[2][########]{%
+  \ifdefined#2%
+    \def\@temp{\def#2#1}%
+    \expandafter\@temp\expandafter{#2}%
+  \else
+    \@latex@error{\noexpand#2undefined}\@ehc
+  \fi
+}
+\@onlypreamble\requiredelimiter
+\makeatother
+
+\newcommand\newdelimitedcommand[2]{
+\expandafter\newcommand\csname #1\endcsname{#2}
+\expandafter\requiredelimiter
+\csname #1 \endcsname
+}
+
+\newdelimitedcommand{toolname}{Tool}
+
+% Print \autoref as "Section X.Y.Z"
+\renewcommand*{\sectionautorefname}{Section}
+\renewcommand*{\subsectionautorefname}{Section}
+\renewcommand*{\subsubsectionautorefname}{Section}
+
+% \circled command to print a colored circle.
+% \circled{1} pretty-prints "(1)"
+% This is useful to refer to labels that are embedded within figures.
+\usepackage{tikz}
+\usetikzlibrary{arrows}
+\usetikzlibrary{shapes}
+\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
+            \node[shape=circle,fill=pairedOneLightBlue,inner sep=1pt] (char) {#1};}}
 
@@
 %% Title information
-\title[Short Title]{Full Title}         %% [Short Title] is optional;
+\title{Scientific Report - Summary of Research\\
+Ambizione: ``Automatized compilation of sequential software to a diverse set of
+hardware accelerators''\\
+Grant number: PZ00P2 168016\\
+Period: 01.05.2018– 31.12.2020\\
+Grantee: Tobias Grosser\\
+\today
+}         %% [Short Title] is optional;
                                         %% when present, will be used in
                                         %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-\subtitle{Subtitle}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
 
@@
 %% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
 
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
+% An abstract should consist of six main sentences:
+%  1. Introduction. In one sentence, what’s the topic?
+%  2. State the problem you tackle.
+%  3. Summarize (in one sentence) why nobody else has adequately answered the research question yet.
+%  4. Explain, in one sentence, how you tackled the research question.
+%  5. In one sentence, how did you go about doing the research that follows from your big idea.
+%  6. As a single sentence, what’s the key impact of your research?
 
-%% Paper note
-%% The \thanks command may be used to create a "paper note" ---
-%% similar to a title note or an author note, but not explicitly
-%% associated with a particular element.  It will appear immediately
-%% above the permission/copyright statement.
-\thanks{with paper note}                %% \thanks is optional
-                                        %% can be repeated if necesary
-                                        %% contents suppressed with 'anonymous'
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-Text of abstract \ldots.
-\end{abstract}
+% (http://www.easterbrook.ca/steve/2010/01/how-to-write-a-scientific-abstract-in-six-easy-steps/)
 
@@
 %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
 %% End of generated code
@@
 %% comma separated list
-\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional
 
-\else
 \settopmatter{printacmref=false} % Removes citation information below abstract
@@
 \maketitle
+\newpage
 \ifx\grammarly\grammarlyon 
@@
 
-\section{Introduction}
-
-Of this super cool article \ldots
-\grosser{A simple comment refering to a particular location}
-
-If there is \authorTwo[some text we want to refer to specifically]{this looks
-particularly interesting} we can refer to this text from within our comments.
-
-And here we only underline \authorTwo[Text]{} that is interesting, but do not
-actually comment.
-
-Now we have a sequence of comments,
-\authorThree{Comment 3}
-\authorFour{Comment 4}
-\authorFive{Comment 5}
-\authorSix{Comment 6}
-which should not introduce overlong white space sequences.
-
-~\\
-Our contributions are:
-\grosser{Always state your contributions explicitly: (A) this makes it
-easy for the reader to understand what novelty is presented, and (B)
-these contributions help you to focus. In particular, the objective of
-the remaining paper should be to support the claims stated here.
-}
-\begin{itemize}
-	\item Contribution 1
-	\item Contribution 2
-	\item Contribution 3
-	\item Contribution 4
-\end{itemize}
-
-\section{Related Work}
-
-\grosser{Related work should always be at the end of the document,
-         as it otherwise becomes an obstacle your reader must
-	 overcome before reaching your idea. For details see:
-	 \url{https://www.microsoft.com/en-us/research/academic-program/write-great-research-paper/}
-}
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
 %% Bibliography
@@
 
-%% Appendix
-\appendix
-\section{Appendix}
+\section{Research Work Conducted}
 
-Text of appendix \ldots
+\section{Deviations}
+
+\section{Important Events}
+
+\section{Contributions}
+
+\section{Research Output}
+
+\subsection{Publications (published)}
+\subsection{Publications (accepted)}
+\subsection{Publications (in preparation)}
 

### 2021-04-21 paper.tex
Add referneces
Add a set of publications
Use a short title

@@
 \ifx\paperversion\paperversioncamera
-  \documentclass[manuscript, nonacm]{acmart}\settopmatter{}
+  \documentclass[manuscript, nonacm, natbib]{acmart}\settopmatter{}
 \fi
@@
                         %% http://ctan.org/pkg/subcaption
+\usepackage{bibentry}
+\nobibliography*
 
@@
 %% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
 %% Citation style
@@
 %% Title information
-\title{Scientific Report - Summary of Research\\
+\title[]{Scientific Report - Summary of Research\\
 Ambizione: ``Automatized compilation of sequential software to a diverse set of
@@
 
-%% Bibliography
-%\bibliography{bibfile}
-
 \section{Research Work Conducted}
@@
 \subsection{Publications (published)}
+\begin{itemize}
+	\item \bibentry{chelini2021mlir}
+	\item \bibentry{10.1145/3437801.3441613}
+	\item \bibentry{grosser2020fast}
+	\item \bibentry{chelini2020automatic}
+	\item \bibentry{khan2020polyhedral}
+	\item \bibentry{schueki2020llhd}
+	\item \bibentry{kurth2020mixed}
+	\item \bibentry{kourtis2020compiling}
+\end{itemize}
 \subsection{Publications (accepted)}
+\subsection{Publications (under review)}
 \subsection{Publications (in preparation)}
+\begin{itemize}
+	\item \bibentry{gysi2020ooc}
+\end{itemize}
+
+%% Bibliography
+\bibliographystyle{plainnat}
+\nobibliography{references}
 

### 2021-04-21 paper.tex
deploy: f3f3a37fddb759f036af733bb2113981a3629546
Add papers under submission
Publications
Increase font size
deploy: 855bc6b4a322191e61cc2a64e993c3a940da4331

@@
 \ifx\paperversion\paperversiondraft
-  \documentclass[manuscript,\review anonymous]{acmart}
+  \documentclass[11pt, manuscript,\review anonymous]{acmart}
 \fi
@@
 \ifx\paperversion\paperversioncamera
-  \documentclass[manuscript, nonacm, natbib]{acmart}\settopmatter{}
+  \documentclass[11pt, manuscript, nonacm, natbib]{acmart}\settopmatter{}
 \fi
@@
 
+The primary goals of this Ambizione research project were proposed as:
+\begin{itemize}
+\item Develop a multi-target heterogeneous compute compiler for C++ and Julia code that generates binaries which effectively exploit
+OpenCL-capable accelerators including GPU, Xeon-Phi, and FPGA.
+\item Research new techniques to scale polyhedral optimization to large kernels and to improve their efficiency to a level that permits
+the effective exploration of a large set of mapping choices.
+\item Derive effective performance models based on hybrid analytical/machine-learning techniques and use them to analyze and optimize accelerator mapping decisions in a large body of existing software.
+\item Evolve our new performance modeling techniques to work within the resource constraints of Julias dynamic run-time system and use Julia to experiment with distributed learning techniques.
+\end{itemize}
+
+The focus of year three in this Ambizione project was move our polyhedral core
+infrastructure into real-world compilers such as MLIR and improve its
+performance. In addition, we worked on establishing backend abstractions for 
+dral techniques. In the
+following, we discuss four milestones we identified as important to reach these goals:
+1. Fast, robust, and (provably?) correct polyhedral libraries
+2. Solid polyhedral compiler infrastructure
+• Modular pattern-based optimizations for CPU and GPU systems
+• DeLICM: Scalar dependence removal
+3. Precise cache models for fully-associative caches
+4. Domain-specific optimization for climate science
+The following two goals from year one remain important:
+1. Solidify and formalize our underlying models
+2. Effective accelerator offloading for large-scale programs
+
 \section{Deviations}
@@
 \end{itemize}
-\subsection{Publications (accepted)}
 \subsection{Publications (under review)}
-\subsection{Publications (in preparation)}
 \begin{itemize}
 	\item \bibentry{gysi2020ooc}
+	\item \bibentry{fehr2021}
+	\item \bibentry{anurudth2021}
+	\item \bibentry{theo2021}
+	\item \bibentry{pichanatan2021}
+	\item \bibentry{copic2021}
 \end{itemize}

### 2021-04-21 paper.tex
Use an enumeratiob
deploy: be4f93ae67ef9288b8f7d0937ee421f3fcbb13ce

@@
 The focus of year three in this Ambizione project was move our polyhedral core
-infrastructure into real-world compilers such as MLIR and improve its
-performance. In addition, we worked on establishing backend abstractions for 
-dral techniques. In the
-following, we discuss four milestones we identified as important to reach these goals:
-1. Fast, robust, and (provably?) correct polyhedral libraries
-2. Solid polyhedral compiler infrastructure
-• Modular pattern-based optimizations for CPU and GPU systems
-• DeLICM: Scalar dependence removal
-3. Precise cache models for fully-associative caches
-4. Domain-specific optimization for climate science
-The following two goals from year one remain important:
-1. Solidify and formalize our underlying models
-2. Effective accelerator offloading for large-scale programs
+infrastructure into real-world compilers (e.g, LLVM/MLIR) and improve its
+performance and scalability. In addition, we worked on establishing backends
+for targetting accelerators such as GPUs and FPGAs. Finally, we worked on
+performance modelling. As mentioned in the year 2 report, our focus moved from
+Julia to MLIR as a compiler and programming language infrastructure.
+
+In the following, we discuss four milestones we identified as important to
+reach these goals:
+
+\begin{enumerate}
+\item Fast, robust, and correct polyhedral libraries
+\item Solid polyhedral compiler infrastructure
+\begin{itemize}
+\item  Modular pattern-based optimizations for CPU and GPU systems
+\item 
+DeLICM: Scalar dependence removal
+\end{itemize}
+\item Domain-specific optimization for climate science
+\end{enumerate}
+
+\subsection{Fast, robust, and correct polyhedral libraries}
+Presburger arithmetic provides the mathematical core for the polyhedral
+compilation techniques that drive analytical cache models, loop optimization
+for ML and HPC, formal verification, and even hardware design.  Polyhedral
+compilation is widely regarded as being slow due to the potentially high
+computational cost of the underlying Presburger libraries. Researchers
+typically use these libraries as powerful black-box tools, but a lack of
+internal documentation and decade-old C implementations hold back broader
+performance-optimization efforts. With FPL, we introduced a new library for
+Presburger arithmetic for the MLIR compiler built from the ground up in modern
+C++. We carefully documented its internal algorithmic foundations, use
+lightweight C++ data structures to minimize memory management costs, and deployed
+transprecision computing across the entire library to effectively exploit
+machine integers and vector instructions. On a newly-developed comprehensive
+benchmark suite for Presburger arithmetic, we show a 5.25x speedup in total
+runtime over the state-of-the-art library isl in its default configuration and
+3.14x when over a variant of isl optimized with element-wise transprecision
+computing. We expect that the availability of a well-documented and fast
+Presburger library will accelerate the adoption of polyhedral compilation
+techniques in production compilers.
+
+The central compnents of this library have already been upstreamed into the
+open-source LLVM/MLIR project and are usable by the wider research and 
+industrial community.
+
+The work in this subproject has been published in \citet{grosser2020fast} and
+\citet{pichanatan2021}.
+
+\subsection{Domain-specific optimization for climate science}
+
+Most compilers have a single core intermediate representation (IR) (e.g., LLVM)
+sometimes complemented with vaguely-defined IR-like data structures. This IR is
+commonly low-level and close to machine instructions. As a result,
+optimizations relying on domain-specific information are either not possible or
+require complex analysis to recover the missing information. In contrast,
+multi-level rewriting instantiates a hierarchy of dialects (IRs), lowers
+programs level-by-level, and performs code transformations at the most suitable
+level. We demonstrate the effectiveness of this approach for the weather and
+climate domain. In particular, we develop a prototype compiler and design
+stencil- and GPU-specific dialects based on a set of newly introduced design
+principles. We find that two domain-specific optimizations (500 lines of code)
+realized on top of LLVM’s extensible MLIR compiler infrastructure suffice to
+outperform state-of-the-art solutions. In essence, multi-level rewriting
+promises to herald the age of specialized compilers composed from domain- and
+target-specific dialects implemented on top of a shared infrastructure.
+
+The GPU backend of the open-earth compiler has already been upstreamed into
+the open-source LLVM/MLIR project and is useable by the wider research
+and industrial community.
+
+The work in this subproject has been published in \citet{gysi2020ooc}
 
@@
 \subsection{Publications (published)}
-\begin{itemize}
+\begin{enumerate}
 	\item \bibentry{chelini2021mlir}
@@
 	\item \bibentry{kourtis2020compiling}
-\end{itemize}
+\end{enumerate}
 \subsection{Publications (under review)}
-\begin{itemize}
+\begin{enumerate}
 	\item \bibentry{gysi2020ooc}
@@
 	\item \bibentry{copic2021}
-\end{itemize}
+\end{enumerate}
 

### 2021-04-21 paper.tex
deploy: 526e7bc303288efc892c17c9db77d3a9497bc084

@@
 
-The central compnents of this library have already been upstreamed into the
+The central components of this library have already been upstreamed into the
 open-source LLVM/MLIR project and are usable by the wider research and 

### 2021-04-21 paper.tex
update title

@@
 %% Title information
-\title[]{Scientific Report - Summary of Research\\
-Ambizione: ``Automatized compilation of sequential software to a diverse set of
-hardware accelerators''\\
+\title[]{
+	{
+	\centering
+	Scientific Report - Summary of Research\\
+	~\\
+Automatized compilation of sequential software to a diverse set of
+hardware accelerators\\
+}
+~\\
+\vspace{18em}
+\large
 Grant number: PZ00P2 168016\\

### 2021-04-21 paper.tex
deploy: 0c2d39d9b19dba8d8383e366560053e5a097a585

@@
 	~\\
-Automatized compilation of sequential software to a diverse set of
-hardware accelerators\\
+``Automatized compilation of sequential software to a diverse set of
+hardware accelerators''\\
 }

### 2021-04-21 paper.tex
WIP

--- 2021-04-21 paper.tex
@@
 
-The focus of year three in this Ambizione project was move our polyhedral core
-infrastructure into real-world compilers (e.g, LLVM/MLIR) and improve its
-performance and scalability. In addition, we worked on establishing backends
-for targetting accelerators such as GPUs and FPGAs. Finally, we worked on
-performance modelling. As mentioned in the year 2 report, our focus moved from
-Julia to MLIR as a compiler and programming language infrastructure.
+The focus of year three in this Ambizione project was the integration of a scalable
+polyhedral core infrastructure into real-world compilers (e.g, LLVM/MLIR),
+establishing backends for targetting accelerators such as GPUs and FPGAs, and,
+finally, developing good performance models. As mentioned in the year 2
+report, our focus moved from Julia to MLIR as a compiler and programming
+language infrastructure.
 
@@
 \end{itemize}
-\item Domain-specific optimization for climate science
+\item Fast GPU code generation for climate models
 \end{enumerate}
@@
 
-\subsection{Domain-specific optimization for climate science}
+\subsection{Fast GPU code generation for climate models}
 

### 2021-04-21 paper.tex
Make report oneside
deploy: 8416320dd7f6741afdc2a132465840fdf0ef1113
WIP

@@
 \ifx\paperversion\paperversioncamera
-  \documentclass[11pt, manuscript, nonacm, natbib]{acmart}\settopmatter{}
+  \documentclass[11pt, manuscript, nonacm, natbib, oneside]{acmart}\settopmatter{}
 \fi
 
+\geometry{twoside=false}
 \usepackage{colortbl}
@@
 \end{itemize}
-\item Fast GPU code generation for climate models
+\item Fast GPU code generation demonstrated via a climate DSL
+\item Code Generation for Emerging Architectures and FPGA
 \end{enumerate}
@@
 open-source LLVM/MLIR project and are usable by the wider research and 
-industrial community.
-
-The work in this subproject has been published in \citet{grosser2020fast} and
+industrial community. The work in this subproject has been published in \citet{grosser2020fast} and
 \citet{pichanatan2021}.
 
-\subsection{Fast GPU code generation for climate models}
+\subsection{Fast GPU code generation demonstrated via a climate DSL}
 
@@
 the open-source LLVM/MLIR project and is useable by the wider research
-and industrial community.
+and industrial community. The work in this subproject has been published in \citet{gysi2020ooc}
 
-The work in this subproject has been published in \citet{gysi2020ooc}
+\subsection{Code Generation for Emerging Architectures and FPGA}
+
+Automatic optimization for emerging architectures and FPGA has been a principal
+objective of this work. We exploited the polyhedral techniques we developed in
+earlier tasks to generate high-performance code for emerging memory
+architectures such as racetrack memories to optimize polyhedral loop nests and
+deep neural networks. We also proposed infrastructure for an open-source
+hardware stack that allows for the compilation and simulation of System Verilog
+code. We also observed an increased use of systems with where the data models
+(e.g., pointer sizes) differ between host systems and accelerators. To ensure
+the automatic code generation for such architectures that also allows for
+automatic data movement, we expanded LLVM and its OpenMP software stack to
+support mixed-data-model offloading.
+
+The LLHD compiler has been upstreamed as open-source into the LLVM/CIRCT
+project. The work in this subproject has been published in \citet{khan2020polyhedral},
+\citet{schueki2020llhd}, \citet{kurth2020mixed}, and \citet{kourtis2020compiling}.
 

### 2021-04-21 paper.tex
Report

@@
 \item Fast, robust, and correct polyhedral libraries
-\item Solid polyhedral compiler infrastructure
-\begin{itemize}
-\item  Modular pattern-based optimizations for CPU and GPU systems
-\item 
-DeLICM: Scalar dependence removal
-\end{itemize}
+\item Automatic Raising to high-level descriptions
 \item Fast GPU code generation demonstrated via a climate DSL
-\item Code Generation for Emerging Architectures and FPGA
+\item Code generation for emerging architectures and FPGA
+\item Scalable performance analysis within a compiler.
 \end{enumerate}
@@
 industrial community. The work in this subproject has been published in \citet{grosser2020fast} and
-\citet{pichanatan2021}.
+is under review in \citet{pichanatan2021}.
+
+\subsection{Automatic Raising to high-level descriptions}
+
+Multi-level intermediate representations (IR) show great promise for lowering
+the design costs for domain-specific compilers by providing a reusable,
+extensible, and non-opinionated framework for expressing domain-specific and
+high-level abstractions directly in the IR. But, while such frameworks support
+the progressive lowering of high-level representations to low-level IR, they do
+not raise in the opposite direction. Thus, the entry point into the compilation
+pipeline defines the highest level of abstraction for all subsequent
+transformations, limiting the set of applicable optimizations, in particular
+for general-purpose languages that are not semantically rich enough to model
+the required abstractions.  We propose Progressive Raising, a complementary
+approach to the progressive lowering in multi-level IRs that raises from lower
+to higher-level abstractions to leverage domain-specific transformations for
+low-level representations. We further introduce MultiLevel Tactics, our
+declarative approach for progressive raising, implemented on top of the MLIR
+framework, and demonstrate the progressive raising from affine loop nests
+specified in a general-purpose language to high-level linear algebra
+operations.  Our raising paths leverage subsequent high-level domain-specific
+transformations with significant performance improvements.
+
+All this work has been has already been upstreamed into the open-source
+LLVM/MLIR project and is usable by the wider research and industrial community.
+The work in this subproject has been published in \citet{chelini2021mlir}.
 
@@
 
+\subsection{Scalable performance analysis within a compiler}
+
+Performance models are well-known instruments to understand the scaling
+behavior of parallel applications. They express how performance changes as key
+execution parameters, such as the number of processes or the size of the input
+problem, vary. Besides reasoning about program behavior, such models can also
+be automatically derived from performance data. This is called empirical
+performance modeling. While this sounds simple at the first glance, this
+approach faces several serious interrelated challenges, including expensive
+performance measurements, inaccuracies inflicted by noisy benchmark data, and
+overall complex experiment design, starting with the selection of the right
+parameters. The more parameters one considers, the more experiments are needed
+and the stronger the impact of noise. In this paper, we show how taint
+analysis, a technique borrowed from the domain of computer security, can
+substantially improve the modeling process, lowering its cost, improving model
+quality, and help validate performance models and experimental setups.
+
+In a subsequent step, we analyzed the optimality of function inlining as a means
+to minimize code size and alter expose further optimization opportunities. This
+work has shown promising code size reductions on SPEC 2017 and the LLVM compiler.
+We hope to use it in the future as an enabling transformation for large-scale
+automatic accelerator mapping.
+
+The work in this subproject has been published in \citet{copic2021} and is under review in
+\citet{theo2021}.
+
 \section{Deviations}

### 2021-04-21 paper.tex
Report
Scientific report

@@
 
+Compared to the original proposal we continued to focus on LLVM/MLIR as an
+infrastructure instead of Julia and not only used it to bring our ideas closer
+to production compilers but indeed contributed many of our ideas directly into
+a widely used software stack.
+
 \section{Important Events}
+
+01.02.2020 - Hiring of Christoph Mueller\\
+01.11.2019 - Hiring of Tobias Gysi
 
@@
 
+Tobias Grosser was leading the overall research agenda and lead the development
+of our novel polyhedral core library. Tobias Gysi and Christoph Mueller have
+been pivotal to the development of the climate DSL and the GPU backend that we
+developed. They did not only contribute to the scientific ideas, but
+contributed high-quality code to the LLVM/MLIR framework directly. As a result,
+they got hired by MeteoSwiss and Google right after the termination of this
+project. Theodoris Theodoridis contributed significantly to both the polyhedral
+math library as well as the inlining model we developed. Tobias Gysi was also
+pivotal in supporting most other work on the LLVM/MLIR side that took place via
+interns and researchers hired via our industrial funding. In particular, in
+2019 visitors played a pivotal role: (1) Arjun Pitchanathan as contributor
+to our fast simplex and the fast Presburger library we built on top, (2)
+Mathieu Fehr as contributor to our declarative compilation work, (3) Lorenzo
+Chelini as a long-term visiting researcher contributing to pattern recognition
+and our declarative compilation infrastructure, (4) Asif Ali Kahn on applying
+our polyhedral tools to racetrack memory.
+
 \section{Research Output}
+
+In the third part of this project, we published eight papers (seven peer
+reviewed) and have six additional papers currently under review.
 

### 2021-04-21 paper.tex
deploy: 0f462f07ae66d814c22a2432d8a25851cde7a270

@@
 
-In the following, we discuss four milestones we identified as important to
+In the following, we discuss five milestones we identified as important to
 reach these goals:
@@
 \end{enumerate}
+
+Overall, this project developed important scientific ideas and made many of
+them available in a production compiler setting via the MLIR project. In
+particular, the LLVM project now has a modern and fast polyhedral library,
+a generic GPU backend, a Verilog compiler. These components have been
+tested in various combinations to accelerate climate science, deep neural
+networks, and a number of general-purpose codes. We believe this project
+made critical advances towards its goal of automating the generation
+of high-performance code.
+
+The PI moved to an associate professor position, the staff on this project
+moved to PhD and industry positions working on projects directly related
+to ideas from this proposal, and three of our interns are taking up a
+new PhD position with the PI himself.
 

### 2021-04-21 tools/create-image.py
remove unnecessary print lines

@@ -23,13 +23,11 @@ def removePNGFiles(directory, tmpOnly = False):
       os.remove(directory + '/' + f)
 
 def createImageOfPaper(path, width = 7, pages = 21):
-    print(path)
     removePNGFiles(os.path.dirname(path), tmpOnly=True)
 
     name = path
 
     cmd = ['convert', '-density', '300', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
-    print(cmd)
     run(cmd)
 
     pageNumber = getPageNumber(path);

### 2021-04-21 tools/create-image.py
Half image resolution to speedup image creation

@@ -27,7 +27,7 @@ def createImageOfPaper(path, width = 7, pages = 21):
 
     name = path
 
-    cmd = ['convert', '-density', '300', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
+    cmd = ['convert', '-density', '150', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
     run(cmd)
 
     pageNumber = getPageNumber(path);

### 2021-04-21 tools/create-image.py
deploy: 3035cc1e75ce9bb760fa9a0b6161bae911d30432

@@ -23,13 +23,11 @@ def removePNGFiles(directory, tmpOnly = False):
       os.remove(directory + '/' + f)
 
 def createImageOfPaper(path, width = 7, pages = 21):
-    print(path)
     removePNGFiles(os.path.dirname(path), tmpOnly=True)
 
     name = path
 
-    cmd = ['convert', '-density', '300', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
-    print(cmd)
+    cmd = ['convert', '-density', '150', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
     run(cmd)
 
     pageNumber = getPageNumber(path);

### 2021-09-03 make_black_and_white.sh
Add script to convert a file to a black and white version

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+gs  -sOutputFile=black_and_white.pdf  -sDEVICE=pdfwrite  -sColorConversionStrategy=Gray  -dProcessColorModel=/DeviceGray  -dCompatibilityLevel=1.4  -dNOPAUSE  -dBATCH  $1
+

### 2022-07-11 tools/create-video.py
Use pdfinfo to get number of pages in pdf
This means we know the number of pages before converting a pdf into a
png, which means we can use the page numbers to later decide the
resolution (dpi) of the generated pdfs.

@@ -7,11 +7,11 @@ import sys
 import shutil
 
 def getPageNumber(name):
-    files = os.listdir(os.path.dirname(name))
-    files = filter(lambda f: f.startswith(os.path.basename(name)), files)
-    files = filter(lambda f: f.endswith('.png'), files)
-    files = list(files)
-    return len(files)
+    cmd = "pdfinfo " + name + " | grep 'Pages' | awk '{print $2}'"
+    res = os.popen(cmd).read().strip()
+    if res == "":
+        return 0
+    return int(res)
 
 def removePNGFiles(directory, tmpOnly = False):
     files = os.listdir(directory)

### 2022-07-12 tools/create-video.py
Create videos from git

@@ -5,6 +5,9 @@ import os
 import math
 import sys
 import shutil
+from git import Repo
+import zipfile
+from subprocess import DEVNULL, STDOUT, run
 
 def getPageNumber(name):
     cmd = "pdfinfo " + name + " | grep 'Pages' | awk '{print $2}'"
@@ -12,6 +15,11 @@ def getPageNumber(name):
     if res == "":
         return 0
     return int(res)
+
+def getPDFSize(name):
+    cmd = "pdfinfo " + name + " | grep 'Page size'"
+    res = os.popen(cmd).read().strip().split(" ")
+    return (int(res[8]), int(res[10]))
 
 def removePNGFiles(directory, tmpOnly = False):
     files = os.listdir(directory)
@@ -28,18 +36,38 @@ def createImageOfPaper(path, width = 7, pages = 21):
 
     name = path
 
-    cmd = ['convert', '-density', '300', path, '-resize', '40%', '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
+    pageNumber = getPageNumber(path);
+
+    if pageNumber == 0:
+        return
+
+    grid = getGrid(pageNumber)
+
+    width = grid[0]
+    pages = grid[0] * grid[1]
+
+    pdfSize = getPDFSize(path)
+
+    targetPage = (3840, 2160)
+
+    targetPdf = (targetPage[0] / grid[0], targetPage[1] / grid[1])
+    densityWidth = 100 / pdfSize[0] * (targetPage[0] / grid[0]) / 100 * 72
+    densityHeight = 100 / pdfSize[1] * (targetPage[1] / grid[1]) / 100 * 72
+
+    density = min(densityWidth, densityHeight)
+    rows = grid[1]
+
+    cmd = ['convert', '-quality', '100', '-density', str(density), path, '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]
     run(cmd)
 
-    pageNumber = getPageNumber(path);
     if pageNumber == 1:
         shutil.copyfile(name + "-tmp.png", name + "-tmp-0.png")
 
     for i in range(pageNumber, pages):
-        cmd = ['convert', '-size', '1020x1320', 'xc:white', name + '-tmp-' + str(i) + '.png']
+        cmd = ['convert', '-size', str(1458 / rows) + "x" + str(density * 10), 'xc:white', name + '-tmp-' + str(i) + '.png']
         run(cmd)
 
-    for row in range(0, math.ceil(pages / width)):
[…]

### 2022-09-06 tools/MLIRLexer.py
add more functionality to lexer

@@ -1,6 +1,7 @@
 from pygments.lexer import RegexLexer, bygroups
 from pygments.token import *
 
+
 class MLIRLexer(RegexLexer):
     name = 'MLIR'
     aliases = ['mlir']
@@ -9,16 +10,29 @@ class MLIRLexer(RegexLexer):
     tokens = {
         'root': [
             (r'//.*?\n', Comment),
-            (r'%[^ ]*', Name.Variable),
+            (r'%[^[ )]*]*', Name.Variable),
             (r'@[^(]*', Name.Function),
-            (r'\^[^(]*', Name.Label),
-            (r'(=)( +)([a-z]+)(\.)([a-z]+)', bygroups(Operator, Text, Name.Namespace, Text, Keyword.Function)),
-            (r'(!)([a-z]+)(\.)([a-z]+)', bygroups(Operator, Name.Namespace, Text, Keyword.Type)),
+            (r'\^[^(:\]]*', Name.Label),
+            (r'(=)( +)([a-z_]+)(\.)([a-z_]+)',
+             bygroups(Operator, Text, Name.Namespace, Text, Keyword.Function)),
+            (r'([a-z_]+)(\.)([a-z_]+)',
+             bygroups(Name.Namespace, Text, Keyword.Function)),
+            (r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
+             bygroups(Operator, Name.Namespace, Text, Keyword.Type)),
+            (r'(!)([a-z0-9]+)', bygroups(Operator, Keyword.Type)),
             (r'(\n|\s)+', Text),
+            (r'([{\[])([a-z_]+)( = +)([a-z0-9">=]+)',
+             bygroups(Text, Name.Attribute, Text, Name.Tag)),
+            (r'(\[)([a-z]+)', bygroups(Text, Name.Variable)),
+            (r'(, +)([a-z_]+)( = +)([a-z0-9">=]+)',
+             bygroups(Text, Name.Attribute, Text, Name.Tag)),
+            (r'(, +)([a-z_]+)( = +)', bygroups(Text, Name.Attribute, Text)),
             (r'[=<>{}:\[\]()*.,!]|x\b', Punctuation),
+            (r'(...)', Text),
             (r'def', Keyword),
         ]
     }
+
 
 # This Lexer is a hack that enables a custom style where only operation names
 # are highlighted and comments as well as dialect names are printed in light
[…]

### 2022-09-06 tools/MLIRLexer.py
clearify rules

@@ -1,5 +1,25 @@
 from pygments.lexer import RegexLexer, bygroups
-from pygments.token import *
+from pygments.token import Name, Keyword, Operator, Comment, Text, Punctuation, Literal
+
+comment_rule = (r'//.*?\n', Comment)
+ssa_value_rule = (r'%[^[ )]*]*', Name.Variable)
+symbol_rule = (r'@[^(]*', Name.Function)
+basic_block_rule = (r'@[^(]*', Name.Function)
+operation_rule = (r'(=)( +)([a-z_]+)(\.)([a-z_]+)',
+                  bygroups(Operator, Text, Name.Namespace, Text,
+                           Keyword.Function))
+non_assign_operation_rule = (r'([a-z_]+)(\.)([a-z_]+)',
+                             bygroups(Name.Namespace, Text, Keyword.Function))
+type_rule = (r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
+             bygroups(Operator, Name.Namespace, Text, Keyword.Type))
+abbrev_type_tule = (r'(!)([a-z0-9]+)', bygroups(Operator, Keyword.Type))
+first_attribute_rule = (r'([{\[])([a-z_]+)( = +)([a-z0-9">=]+)',
+                        bygroups(Text, Name.Attribute, Text, Name.Tag))
+following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=]+)',
+                            bygroups(Text, Name.Attribute, Text, Name.Tag))
+abbrev_following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=]+)',
+                                   bygroups(Text, Name.Attribute, Text,
+                                            Name.Tag))
 
 
 class MLIRLexer(RegexLexer):
@@ -9,24 +29,19 @@ class MLIRLexer(RegexLexer):
 
     tokens = {
         'root': [
-            (r'//.*?\n', Comment),
-            (r'%[^[ )]*]*', Name.Variable),
-            (r'@[^(]*', Name.Function),
-            (r'\^[^(:\]]*', Name.Label),
-            (r'(=)( +)([a-z_]+)(\.)([a-z_]+)',
-             bygroups(Operator, Text, Name.Namespace, Text, Keyword.Function)),
-            (r'([a-z_]+)(\.)([a-z_]+)',
-             bygroups(Name.Namespace, Text, Keyword.Function)),
-            (r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
[…]

### 2022-09-06 tools/MLIRLexer.py
fix basic_block_rule

@@ -4,7 +4,7 @@ from pygments.token import Name, Keyword, Operator, Comment, Text, Punctuation, 
 comment_rule = (r'//.*?\n', Comment)
 ssa_value_rule = (r'%[^[ )]*]*', Name.Variable)
 symbol_rule = (r'@[^(]*', Name.Function)
-basic_block_rule = (r'@[^(]*', Name.Function)
+basic_block_rule = (r'\^[^(:\]]*', Name.Label)
 operation_rule = (r'(=)( +)([a-z_]+)(\.)([a-z_]+)',
                   bygroups(Operator, Text, Name.Namespace, Text,
                            Keyword.Function))

### 2022-09-07 tools/MLIRLexer.py
add example

@@ -13,13 +13,12 @@ non_assign_operation_rule = (r'([a-z_]+)(\.)([a-z_]+)',
 type_rule = (r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
              bygroups(Operator, Name.Namespace, Text, Keyword.Type))
 abbrev_type_tule = (r'(!)([a-z0-9]+)', bygroups(Operator, Keyword.Type))
-first_attribute_rule = (r'([{\[])([a-z_]+)( = +)([a-z0-9">=]+)',
+first_attribute_rule = (r'([{\[])([a-z_A-Z]+)( = +)([@a-z0-9">=]+)',
                         bygroups(Text, Name.Attribute, Text, Name.Tag))
-following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=]+)',
+following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=@]+)',
                             bygroups(Text, Name.Attribute, Text, Name.Tag))
-abbrev_following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=]+)',
-                                   bygroups(Text, Name.Attribute, Text,
-                                            Name.Tag))
+abbrev_following_attribute_rule = (r'(, +)([a-z_]+)( = +)',
+                                   bygroups(Text, Name.Attribute, Text))
 
 
 class MLIRLexer(RegexLexer):

### 2022-09-08 tools/MLIRLexer.py
FEAT: more functionality for highlighting MLIR/xDSL Code (#27)

@@ -1,5 +1,25 @@
 from pygments.lexer import RegexLexer, bygroups
-from pygments.token import *
+from pygments.token import Name, Keyword, Operator, Comment, Text, Punctuation, Literal
+
+comment_rule = (r'//.*?\n', Comment)
+ssa_value_rule = (r'%[^[ )]*]*', Name.Variable)
+symbol_rule = (r'@[^(]*', Name.Function)
+basic_block_rule = (r'\^[^(:\]]*', Name.Label)
+operation_rule = (r'(=)( +)([a-z_]+)(\.)([a-z_]+)',
+                  bygroups(Operator, Text, Name.Namespace, Text,
+                           Keyword.Function))
+non_assign_operation_rule = (r'([a-z_]+)(\.)([a-z_]+)',
+                             bygroups(Name.Namespace, Text, Keyword.Function))
+type_rule = (r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
+             bygroups(Operator, Name.Namespace, Text, Keyword.Type))
+abbrev_type_tule = (r'(!)([a-z0-9]+)', bygroups(Operator, Keyword.Type))
+first_attribute_rule = (r'([{\[])([a-z_A-Z]+)( = +)([@a-z0-9">=]+)',
+                        bygroups(Text, Name.Attribute, Text, Name.Tag))
+following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=@]+)',
+                            bygroups(Text, Name.Attribute, Text, Name.Tag))
+abbrev_following_attribute_rule = (r'(, +)([a-z_]+)( = +)',
+                                   bygroups(Text, Name.Attribute, Text))
+
 
 class MLIRLexer(RegexLexer):
     name = 'MLIR'
@@ -8,17 +28,25 @@ class MLIRLexer(RegexLexer):
 
     tokens = {
         'root': [
-            (r'//.*?\n', Comment),
-            (r'%[^ ]*', Name.Variable),
-            (r'@[^(]*', Name.Function),
-            (r'\^[^(]*', Name.Label),
-            (r'(=)( +)([a-z]+)(\.)([a-z]+)', bygroups(Operator, Text, Name.Namespace, Text, Keyword.Function)),
-            (r'(!)([a-z]+)(\.)([a-z]+)', bygroups(Operator, Name.Namespace, Text, Keyword.Type)),
+            comment_rule,
+            ssa_value_rule,
+            symbol_rule,
+            basic_block_rule,
+            operation_rule,
[…]

### 2022-10-21 plots/speedup.py
Make creation date of matplotlib figure deterministic

@@ -72,5 +72,9 @@ autolabel(rects2)
 fig.tight_layout()
 
 filename = os.path.basename(__file__).replace(".py", ".pdf")
-fig.savefig(filename)
 
+# Do not emit a creation date. This will make the content of
+# the pdfs we generate more deterministic.
+metadata = {'CreationDate' : None }
+
+fig.savefig(filename, metadata = metadata)

### 2022-11-12 tools/create-video.py
Support non-integral document sizes

@@ -19,7 +19,7 @@ def getPageNumber(name):
 def getPDFSize(name):
     cmd = "pdfinfo " + name + " | grep 'Page size'"
     res = os.popen(cmd).read().strip().split(" ")
-    return (int(res[8]), int(res[10]))
+    return (int(round(float(res[8]),0)), int(round(float(res[10]),0)))
 
 def removePNGFiles(directory, tmpOnly = False):
     files = os.listdir(directory)

### 2022-11-12 tools/create-video.py
Build image files in parallel

@@ -8,6 +8,7 @@ import shutil
 from git import Repo
 import zipfile
 from subprocess import DEVNULL, STDOUT, run
+import concurrent.futures
 
 def getPageNumber(name):
     cmd = "pdfinfo " + name + " | grep 'Pages' | awk '{print $2}'"
@@ -120,22 +121,29 @@ def createVideo(path = ""):
     run(cmd)
     print(path + 'video.mp4')
 
+def createImage(data):
+    repo = Repo(".")
+    commit_number = data[0]
+    commit = data[1]
+    filename = str(commit_number) + ".zip"
+    dirname =  str(commit_number).zfill(5) + "-files"
+
+    repo.git.archive(commit, '-o', filename)
+    with zipfile.ZipFile(filename, 'r') as zip_ref:
+        zip_ref.extractall(dirname)
+
+    shutil.copyfile("Makefile", dirname + "/Makefile")
+    shutil.copyfile(".latexmkrc", dirname + "/.latexmkrc")
+    run(['make', '-C', dirname, "paper.pdf"], stdout=DEVNULL, stderr=STDOUT)
+    createImageOfPaper(dirname + "/paper.pdf")
+
 def createImages():
     repo = Repo(".")
-    commit_number = 0
-    for commit in reversed(list(repo.iter_commits('master'))):
-        filename = str(commit_number) + ".zip"
-        dirname =  str(commit_number).zfill(5) + "-files"
-        commit_number += 1
+    commits = list(reversed(list(repo.iter_commits('master'))))
+    items = zip(range(len(commits)), commits)
 
-        repo.git.archive(commit, '-o', filename)
-        with zipfile.ZipFile(filename, 'r') as zip_ref:
-            zip_ref.extractall(dirname)
-
-        shutil.copyfile("Makefile", dirname + "/Makefile")
-        shutil.copyfile(".latexmkrc", dirname + "/.latexmkrc")
-        run(['make', '-C', dirname, "paper.pdf"], stdout=DEVNULL, stderr=STDOUT)
-        createImageOfPaper(dirname + "/paper.pdf")
+    with concurrent.futures.ThreadPoolExecutor() as executor:
+        executor.map(createImage, items)
 
 def getGrid(pages):
     import itertools
@@ -167,5 +175,4 @@ def testImages():
         testImageSize(i)
 
 createImages()
-moveImages(imagePath)
-createVideo()
+#createVideo()

### 2022-11-12 tools/create-video.py
create-video: add some useful command line arguments

@@ -9,6 +9,7 @@ from git import Repo
 import zipfile
 from subprocess import DEVNULL, STDOUT, run
 import concurrent.futures
+import argparse
 
 def getPageNumber(name):
     cmd = "pdfinfo " + name + " | grep 'Pages' | awk '{print $2}'"
@@ -137,9 +138,9 @@ def createImage(data):
     run(['make', '-C', dirname, "paper.pdf"], stdout=DEVNULL, stderr=STDOUT)
     createImageOfPaper(dirname + "/paper.pdf")
 
-def createImages():
+def createImages(branch, count):
     repo = Repo(".")
-    commits = list(reversed(list(repo.iter_commits('master'))))
+    commits = list(reversed(list(repo.iter_commits(branch, max_count=count))))
     items = zip(range(len(commits)), commits)
 
     with concurrent.futures.ThreadPoolExecutor() as executor:
@@ -174,5 +175,14 @@ def testImages():
     for i in range(1, 27):
         testImageSize(i)
 
-createImages()
-#createVideo()
+parser = argparse.ArgumentParser(
+    prog = 'ltxrepo2mpg',
+    description = 'Translate a git paper repository into a video')
+parser.add_argument('-c', '--count', default=10000)
+parser.add_argument('-b', '--branch', default='main')
+
+args = parser.parse_args()
+
+
+createImages(args.branch, args.count)
+createVideo()

### 2022-11-12 tools/create-video.py
create-video: plot statistics

@@ -10,6 +10,13 @@ import zipfile
 from subprocess import DEVNULL, STDOUT, run
 import concurrent.futures
 import argparse
+import numpy as np
+import matplotlib.pyplot as plt
+import time
+from datetime import date, timedelta
+from matplotlib.ticker import ScalarFormatter
+import matplotlib.dates as mdates
+
 
 def getPageNumber(name):
     cmd = "pdfinfo " + name + " | grep 'Pages' | awk '{print $2}'"
@@ -33,7 +40,6 @@ def removePNGFiles(directory, tmpOnly = False):
       os.remove(directory + '/' + f)
 
 def createImageOfPaper(path, width = 7, pages = 21):
-    print(path)
     removePNGFiles(os.path.dirname(path))
 
     name = path
@@ -122,7 +128,46 @@ def createVideo(path = ""):
     run(cmd)
     print(path + 'video.mp4')
 
-def createImage(data):
+def plotStatistics(commit, filename, branch, count):
+    repo = Repo(".")
+    commit_number = 0
+    commits = list(repo.iter_commits(branch))
+    end = date.fromtimestamp(commits[0].committed_date)
+    start = date.fromtimestamp(commits[-1].committed_date)
+    delta = end - start
+
+    date_ticks = []
+    for offset in range(delta.days+1):
+        current = start + timedelta(days=offset)
+        date_ticks.append(current)
+
+    histogram = [0] * (delta.days+1)
+
+    for commit in commits:
+      commit_date = date.fromtimestamp(commit.committed_date)
+      histogram[(commit_date - start).days] += 1
+
+    plt.bar(date_ticks, histogram, align='center', alpha=0.5)
+
+    commit_date = date.fromtimestamp(commit.committed_date)
+    delta = end - commit_date
+    histogram_highlight = [0] * (delta.days+1)
+    histogram_highlight[delta.days] = histogram[delta.days]
+
+    plt.bar(date_ticks, histogram_highlight, align='center', alpha=0.5, color='red')
+    ax = plt.gca()
+
+    ax.set_yscale('log')
+    sf = ScalarFormatter()
+    sf.set_scientific(False)
+    ax.yaxis.set_major_formatter(sf)
+    ax.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
+    ax.spines["right"].set_visible(False)
+    ax.spines["top"].set_visible(False)
+    plt.savefig(filename)
+
+def createImage(data, branch, count):
+    print(data)
[…]

### 2022-11-12 tools/create-video.py
create_video: add statistics into video

@@ -147,14 +147,16 @@ def plotStatistics(commit, filename, branch, count):
       commit_date = date.fromtimestamp(commit.committed_date)
       histogram[(commit_date - start).days] += 1
 
-    plt.bar(date_ticks, histogram, align='center', alpha=0.5)
+    fig, ax = plt.subplots(figsize=(5, 3))
+
+    ax.bar(date_ticks, histogram, align='center', alpha=0.5)
 
     commit_date = date.fromtimestamp(commit.committed_date)
     delta = end - commit_date
     histogram_highlight = [0] * (delta.days+1)
     histogram_highlight[delta.days] = histogram[delta.days]
 
-    plt.bar(date_ticks, histogram_highlight, align='center', alpha=0.5, color='red')
+    ax.bar(date_ticks, histogram_highlight, align='center', alpha=0.5, color='red')
     ax = plt.gca()
 
     ax.set_yscale('log')
@@ -164,7 +166,7 @@ def plotStatistics(commit, filename, branch, count):
     ax.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
     ax.spines["right"].set_visible(False)
     ax.spines["top"].set_visible(False)
-    plt.savefig(filename)
+    plt.savefig(filename, dpi=300, transparent=True)
 
 def createImage(data, branch, count):
     print(data)
@@ -183,6 +185,9 @@ def createImage(data, branch, count):
     run(['make', '-C', dirname, "paper.pdf"], stdout=DEVNULL, stderr=STDOUT)
     createImageOfPaper(dirname + "/paper.pdf")
     plotStatistics(commit,  dirname + "/statistics.png", branch, count)
+    run(['convert', '-gravity', 'SouthEast', dirname + '/paper.pdf-full.png',
+        dirname + '/statistics.png', '-composite', dirname +
+        '/paper.pdf-full.png'])
 
 def createImages(branch, count):
     repo = Repo(".")

### 2022-11-12 tools/create-video.py
create_video: run in separate processes
This avoids matplotlib inconsistencies

@@ -16,6 +16,7 @@ import time
 from datetime import date, timedelta
 from matplotlib.ticker import ScalarFormatter
 import matplotlib.dates as mdates
+from multiprocessing import Pool
 
 
 def getPageNumber(name):
@@ -104,14 +105,6 @@ def getReleases(path):
     releases = list(releases)
     return list(releases)
 
-from multiprocessing import Pool
-
-def createImages(path):
-    releases = getReleases(path)
-
-    with Pool(4) as p:
-        p.map(createImageOfPaper, releases)
-
 def moveImages(path):
     removePNGFiles(path)
     releases = getReleases(path)
@@ -125,6 +118,7 @@ def moveImages(path):
 def createVideo(path = ""):
     run(['rm', path + 'video.mp4'])
     cmd = ['ffmpeg', '-r', '2', '-pattern_type', 'glob', '-i', path + '*-files/paper.pdf-full.png', '-vcodec', 'libx264', '-crf', '25', '-pix_fmt', 'yuv420p', '-vf', 'scale=3840:2160:force_original_aspect_ratio=1,pad=3840:2160:(ow-iw)/2:(oh-ih)/2:white,format=rgb24', path + 'video.mp4']
+    print(" ".join(cmd))
     run(cmd)
     print(path + 'video.mp4')
 
@@ -168,11 +162,13 @@ def plotStatistics(commit, filename, branch, count):
     ax.spines["top"].set_visible(False)
     plt.savefig(filename, dpi=300, transparent=True)
 
-def createImage(data, branch, count):
+def createImage(data):
     print(data)
     repo = Repo(".")
     commit_number = data[0]
-    commit = data[1]
+    commit = repo.commit(data[1])
+    branch = data[2]
+    count = data[3]
     filename = str(commit_number) + ".zip"
     dirname =  str(commit_number).zfill(5) + "-files"
 
@@ -192,13 +188,11 @@ def createImage(data, branch, count):
 def createImages(branch, count):
     repo = Repo(".")
     commits = list(reversed(list(repo.iter_commits(branch, max_count=count))))
-    items = zip(range(len(commits)), commits)
+    commits = list(map(lambda x: x.hexsha, commits))
+    items = zip(range(len(commits)), commits, [branch] * len(commits), [count] * len(commits))
 
-    def f(data):
-        createImage(data, branch, count)
-
-    with concurrent.futures.ThreadPoolExecutor() as executor:
-        executor.map(f, items)
+    with Pool(16) as p:
[…]

### 2022-11-12 tools/create-video.py
create_video: add commit information

@@ -141,7 +141,7 @@ def plotStatistics(commit, filename, branch, count):
       commit_date = date.fromtimestamp(commit.committed_date)
       histogram[(commit_date - start).days] += 1
 
-    fig, ax = plt.subplots(figsize=(5, 3))
+    fig, ax = plt.subplots(figsize=(7, 3))
 
     ax.bar(date_ticks, histogram, align='center', alpha=0.5)
 
@@ -160,6 +160,9 @@ def plotStatistics(commit, filename, branch, count):
     ax.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
     ax.spines["right"].set_visible(False)
     ax.spines["top"].set_visible(False)
+    ax.set_title(commit.hexsha[:12], loc='left')
+    ax.set_title(commit.author.name, loc='center')
+    ax.set_title(time.strftime("%a, %d %b %H:%M", time.gmtime(commit.committed_date)), loc='right')
     plt.savefig(filename, dpi=300, transparent=True)
 
 def createImage(data):

### 2022-11-12 tools/create-video.py
Update commits properly

@@ -137,8 +137,8 @@ def plotStatistics(commit, filename, branch, count):
 
     histogram = [0] * (delta.days+1)
 
-    for commit in commits:
-      commit_date = date.fromtimestamp(commit.committed_date)
+    for c in commits:
+      commit_date = date.fromtimestamp(c.committed_date)
       histogram[(commit_date - start).days] += 1
 
     fig, ax = plt.subplots(figsize=(7, 3))

### 2022-11-12 tools/create-video.py
create_video: fix highlight

@@ -146,11 +146,11 @@ def plotStatistics(commit, filename, branch, count):
     ax.bar(date_ticks, histogram, align='center', alpha=0.5)
 
     commit_date = date.fromtimestamp(commit.committed_date)
-    delta = end - commit_date
-    histogram_highlight = [0] * (delta.days+1)
+    delta = commit_date - start
+    histogram_highlight = [0] * len(histogram)
     histogram_highlight[delta.days] = histogram[delta.days]
 
-    ax.bar(date_ticks, histogram_highlight, align='center', alpha=0.5, color='red')
+    ax.bar(date_ticks, histogram_highlight, align='center', alpha=1, color='red')
     ax = plt.gca()
 
     ax.set_yscale('log')
@@ -160,9 +160,9 @@ def plotStatistics(commit, filename, branch, count):
     ax.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
     ax.spines["right"].set_visible(False)
     ax.spines["top"].set_visible(False)
-    ax.set_title(commit.hexsha[:12], loc='left')
+    ax.set_title(time.strftime("%a, %d %b %H:%M", time.gmtime(commit.committed_date)), loc='left')
     ax.set_title(commit.author.name, loc='center')
-    ax.set_title(time.strftime("%a, %d %b %H:%M", time.gmtime(commit.committed_date)), loc='right')
+    ax.set_title(commit.hexsha[:12], loc='right')
     plt.savefig(filename, dpi=300, transparent=True)
 
 def createImage(data):

### 2022-11-12 tools/create-video.py
create_video: add progress bar

@@ -17,6 +17,7 @@ from datetime import date, timedelta
 from matplotlib.ticker import ScalarFormatter
 import matplotlib.dates as mdates
 from multiprocessing import Pool
+import tqdm
 
 
 def getPageNumber(name):
@@ -192,10 +193,11 @@ def createImages(branch, count):
     repo = Repo(".")
     commits = list(reversed(list(repo.iter_commits(branch, max_count=count))))
     commits = list(map(lambda x: x.hexsha, commits))
-    items = zip(range(len(commits)), commits, [branch] * len(commits), [count] * len(commits))
+    items = list(zip(range(len(commits)), commits, [branch] * len(commits), [count] * len(commits)))
 
-    with Pool(16) as p:
-        p.map(createImage, items)
+    p = Pool(16)
+    for _ in tqdm.tqdm(p.imap_unordered(createImage, items), total=len(items)):
+        pass
 
 def getGrid(pages):
     import itertools

### 2022-11-12 tools/create-video.py
Remove verbose noise

@@ -119,7 +119,6 @@ def moveImages(path):
 def createVideo(path = ""):
     run(['rm', path + 'video.mp4'])
     cmd = ['ffmpeg', '-r', '2', '-pattern_type', 'glob', '-i', path + '*-files/paper.pdf-full.png', '-vcodec', 'libx264', '-crf', '25', '-pix_fmt', 'yuv420p', '-vf', 'scale=3840:2160:force_original_aspect_ratio=1,pad=3840:2160:(ow-iw)/2:(oh-ih)/2:white,format=rgb24', path + 'video.mp4']
-    print(" ".join(cmd))
     run(cmd)
     print(path + 'video.mp4')
 
@@ -167,7 +166,6 @@ def plotStatistics(commit, filename, branch, count):
     plt.savefig(filename, dpi=300, transparent=True)
 
 def createImage(data):
-    print(data)
     repo = Repo(".")
     commit_number = data[0]
     commit = repo.commit(data[1])

### 2022-11-12 tools/create-video.py
create_video: polish figures

@@ -118,7 +118,7 @@ def moveImages(path):
 
 def createVideo(path = ""):
     run(['rm', path + 'video.mp4'])
-    cmd = ['ffmpeg', '-r', '2', '-pattern_type', 'glob', '-i', path + '*-files/paper.pdf-full.png', '-vcodec', 'libx264', '-crf', '25', '-pix_fmt', 'yuv420p', '-vf', 'scale=3840:2160:force_original_aspect_ratio=1,pad=3840:2160:(ow-iw)/2:(oh-ih)/2:white,format=rgb24', path + 'video.mp4']
+    cmd = ['ffmpeg', '-r', '2', '-pattern_type', 'glob', '-i', path + '*-files/paper.pdf-overlay.png', '-vcodec', 'libx264', '-crf', '25', '-pix_fmt', 'yuv420p', '-vf', 'scale=3840:2160:force_original_aspect_ratio=1,pad=3840:2160:(ow-iw)/2:(oh-ih)/2:white,format=rgb24', path + 'video.mp4']
     run(cmd)
     print(path + 'video.mp4')
 
@@ -137,11 +137,19 @@ def plotStatistics(commit, filename, branch, count):
 
     histogram = [0] * (delta.days+1)
 
+
+    commit_pos = 0
+    seen = False
     for c in commits:
+      if not seen:
+          commit_pos += 1
+          if c == commit:
+              seen = True
+      
       commit_date = date.fromtimestamp(c.committed_date)
       histogram[(commit_date - start).days] += 1
 
-    fig, ax = plt.subplots(figsize=(7, 3))
+    fig, (title, ax) = plt.subplots(1, 2, figsize=(10, 1.5))
 
     ax.bar(date_ticks, histogram, align='center', alpha=0.5)
 
@@ -158,11 +166,19 @@ def plotStatistics(commit, filename, branch, count):
     sf.set_scientific(False)
     ax.yaxis.set_major_formatter(sf)
     ax.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
-    ax.spines["right"].set_visible(False)
+    ax.spines["left"].set_visible(False)
     ax.spines["top"].set_visible(False)
-    ax.set_title(time.strftime("%a, %d %b %H:%M", time.gmtime(commit.committed_date)), loc='left')
-    ax.set_title(commit.author.name, loc='center')
-    ax.set_title(commit.hexsha[:12], loc='right')
+    title.spines["left"].set_visible(False)
+    title.spines["top"].set_visible(False)
[…]

### 2022-11-12 tools/create-video.py
create_video: make things work again

@@ -196,8 +196,8 @@ def createImage(data):
 
     shutil.copyfile("Makefile", dirname + "/Makefile")
     shutil.copyfile(".latexmkrc", dirname + "/.latexmkrc")
-    #run(['make', '-C', dirname, "paper.pdf"], stdout=DEVNULL, stderr=STDOUT)
-    #createImageOfPaper(dirname + "/paper.pdf")
+    run(['make', '-C', dirname, "paper.pdf"], stdout=DEVNULL, stderr=STDOUT)
+    createImageOfPaper(dirname + "/paper.pdf")
     plotStatistics(commit,  dirname + "/statistics.png", branch, count)
     run(['convert', dirname + '/paper.pdf-full.png', '-resize', '3840x2160',
         '-background', 'white', '-gravity', 'center', '-extent', '3840x2160',

### 2022-11-12 tools/create-video.py
create_video: fix commit count

@@ -175,7 +175,7 @@ def plotStatistics(commit, filename, branch, count):
     ax.yaxis.tick_right()
     title.text(0, 0, time.strftime("%a, %d %b %H:%M", time.gmtime(commit.committed_date)), horizontalalignment="left", verticalalignment="top")
     title.text(0.5, 0, commit.author.name, horizontalalignment="center", verticalalignment="top")
-    title.text(1, 0, str(commit_pos) + "/" + str(count) + "         ", horizontalalignment="right", verticalalignment="top")
+    title.text(1, 0, str(commit_pos) + "/" + str(min(count, len(commits))) + "         ", horizontalalignment="right", verticalalignment="top")
     title.get_xaxis().set_visible(False)
     title.get_yaxis().set_visible(False)
     plt.tight_layout()

### 2022-11-12 tools/create-video.py
Count commits correctly

@@ -140,7 +140,7 @@ def plotStatistics(commit, filename, branch, count):
 
     commit_pos = 0
     seen = False
-    for c in commits:
+    for c in reverse(commits):
       if not seen:
           commit_pos += 1
           if c == commit:

### 2022-11-12 tools/create-video.py
Count commits correctly

@@ -140,7 +140,9 @@ def plotStatistics(commit, filename, branch, count):
 
     commit_pos = 0
     seen = False
-    for c in reverse(commits):
+    rev_commits = commits
+    rev_commits.reverse()
+    for c in rev_commits:
       if not seen:
           commit_pos += 1
           if c == commit:

### 2022-11-12 tools/create-video.py
Count commits correctly

@@ -140,7 +140,7 @@ def plotStatistics(commit, filename, branch, count):
 
     commit_pos = 0
     seen = False
-    rev_commits = commits
+    rev_commits = list(commits)
     rev_commits.reverse()
     for c in rev_commits:
       if not seen:

### 2022-11-12 tools/create-video.py
create_video: fix assert in -c runs

@@ -177,7 +177,8 @@ def plotStatistics(commit, filename, branch, count):
     ax.yaxis.tick_right()
     title.text(0, 0, time.strftime("%a, %d %b %H:%M", time.gmtime(commit.committed_date)), horizontalalignment="left", verticalalignment="top")
     title.text(0.5, 0, commit.author.name, horizontalalignment="center", verticalalignment="top")
-    title.text(1, 0, str(commit_pos) + "/" + str(min(count, len(commits))) + "         ", horizontalalignment="right", verticalalignment="top")
+    commits_to_visit = min(int(count), len(commits))
+    title.text(1, 0, str(commit_pos) + "/" + str(commits_to_visit) + "         ", horizontalalignment="right", verticalalignment="top")
     title.get_xaxis().set_visible(False)
     title.get_yaxis().set_visible(False)
     plt.tight_layout()

### 2022-11-12 tools/create-video.py
- made a bit more OS-agnostic by using PyPDF2 instead of pdfinfo (so it runs on mac) - added arguments to control num of processes and base filename of paper

@@ -17,20 +17,19 @@ from datetime import date, timedelta
 from matplotlib.ticker import ScalarFormatter
 import matplotlib.dates as mdates
 from multiprocessing import Pool
+import PyPDF2
 import tqdm
 
 
-def getPageNumber(name):
-    cmd = "pdfinfo " + name + " | grep 'Pages' | awk '{print $2}'"
-    res = os.popen(cmd).read().strip()
-    if res == "":
-        return 0
-    return int(res)
-
 def getPDFSize(name):
-    cmd = "pdfinfo " + name + " | grep 'Page size'"
-    res = os.popen(cmd).read().strip().split(" ")
-    return (int(round(float(res[8]),0)), int(round(float(res[10]),0)))
+    with open(name, 'rb') as f:
+        pdf = PyPDF2.PdfFileReader(f)
+        return (pdf.pages[0].mediabox.width,pdf.pages[0].mediabox.height)
+
+def getPageNumber(name):
+    with open(name, 'rb') as f:
+        pdf = PyPDF2.PdfFileReader(f)
+        return pdf.numPages
 
 def removePNGFiles(directory, tmpOnly = False):
     files = os.listdir(directory)
@@ -45,6 +44,8 @@ def createImageOfPaper(path, width = 7, pages = 21):
     removePNGFiles(os.path.dirname(path))
 
     name = path
+    if not os.path.exists(path):
+        return
 
     pageNumber = getPageNumber(path);
 
@@ -116,9 +117,9 @@ def moveImages(path):
         shutil.copyfile(release, path + ("still-%05d.png" % idx))
         idx = idx + 1
 
-def createVideo(path = ""):
+def createVideo(path = "", basename = "paper"):
     run(['rm', path + 'video.mp4'])
-    cmd = ['ffmpeg', '-r', '2', '-pattern_type', 'glob', '-i', path + '*-files/paper.pdf-overlay.png', '-vcodec', 'libx264', '-crf', '25', '-pix_fmt', 'yuv420p', '-vf', 'scale=3840:2160:force_original_aspect_ratio=1,pad=3840:2160:(ow-iw)/2:(oh-ih)/2:white,format=rgb24', path + 'video.mp4']
[…]

### 2022-11-13 plots/speedup.py
plotting: move global settings into a function

@@ -5,11 +5,16 @@ import matplotlib
 import matplotlib.pyplot as plt
 import numpy as np
 
-# Use TrueType fonts instead of Type 3 fonts
-# Type 3 fonts embed bitmaps and are not allowed in camera-ready submissions
-# for many conferences. TrueType fonts look better and are accepted.
-matplotlib.rcParams['pdf.fonttype'] = 42
-matplotlib.rcParams['ps.fonttype'] = 42
+def setGlobalDefaults():
+    ## Use TrueType fonts instead of Type 3 fonts
+    #
+    # Type 3 fonts embed bitmaps and are not allowed in camera-ready submissions
+    # for many conferences. TrueType fonts look better and are accepted.
+    # This follows: https://www.conference-publishing.com/Help.php
+    matplotlib.rcParams['pdf.fonttype'] = 42
+    matplotlib.rcParams['ps.fonttype'] = 42
+
+setGlobalDefaults()
 
 matplotlib.rcParams['figure.figsize'] = 5, 2
 

### 2022-11-13 plots/plot.py
plotting: format file with yapf

@@ -5,6 +5,7 @@ import matplotlib
 import matplotlib.pyplot as plt
 import numpy as np
 
+
 def setGlobalDefaults():
     ## Use TrueType fonts instead of Type 3 fonts
     #
@@ -13,6 +14,7 @@ def setGlobalDefaults():
     # This follows: https://www.conference-publishing.com/Help.php
     matplotlib.rcParams['pdf.fonttype'] = 42
     matplotlib.rcParams['ps.fonttype'] = 42
+
 
 setGlobalDefaults()
 
@@ -36,20 +38,30 @@ x = np.arange(len(labels))  # the label locations
 width = 0.35  # the width of the bars
 
 fig, ax = plt.subplots()
-rects1 = ax.bar(x - width/2, men_means, width, label='Men', color = light_blue)
-rects2 = ax.bar(x + width/2, women_means, width, label='Women', color = dark_blue)
+rects1 = ax.bar(x - width / 2, men_means, width, label='Men', color=light_blue)
+rects2 = ax.bar(x + width / 2,
+                women_means,
+                width,
+                label='Women',
+                color=dark_blue)
 
 # Y-Axis Label
 #
 # Use a horizontal label for improved readability.
-ax.set_ylabel('Speedup', rotation='horizontal', position = (1, 1.05),
-    horizontalalignment='left', verticalalignment='bottom')
+ax.set_ylabel('Speedup',
+              rotation='horizontal',
+              position=(1, 1.05),
+              horizontalalignment='left',
+              verticalalignment='bottom')
 
 # Add some text for labels, title and custom x-axis tick labels, etc.
 ax.set_xticks(x)
 ax.set_xticklabels(labels)
 
-ax.legend(ncol=100, frameon=False, loc='lower right', bbox_to_anchor=(0, 1, 1, 0))
+ax.legend(ncol=100,
+          frameon=False,
+          loc='lower right',
+          bbox_to_anchor=(0, 1, 1, 0))
 
 # Hide the right and top spines
 #
@@ -60,16 +72,20 @@ ax.legend(ncol=100, frameon=False, loc='lower right', bbox_to_anchor=(0, 1, 1, 0
 ax.spines['right'].set_visible(False)
 ax.spines['top'].set_visible(False)
 
+
 def autolabel(rects):
     """Attach a text label above each bar in *rects*, displaying its height."""
     for rect in rects:
         height = rect.get_height()
-        ax.annotate('{}'.format(height),
-                    xy=(rect.get_x() + rect.get_width() / 2, height),
[…]

### 2022-11-13 plots/plot.py
plotting: format imports

@@ -1,6 +1,7 @@
 #!/usr/bin/env python3
 
 import os
+
 import matplotlib
 import matplotlib.pyplot as plt
 import numpy as np

### 2022-11-13 plots/plot.py
plotting: add argument parsing to only print one plot
This will make it easier to selectively update plots in a paper.

@@ -1,5 +1,6 @@
 #!/usr/bin/env python3
 
+import argparse
 import os
 
 import matplotlib
@@ -17,13 +18,7 @@ def setGlobalDefaults():
     matplotlib.rcParams['ps.fonttype'] = 42
 
 
-setGlobalDefaults()
-
 matplotlib.rcParams['figure.figsize'] = 5, 2
-
-labels = ['G1', 'G2', 'G3', 'G4', 'G5']
-men_means = [1.5, 1.2, 1.3, 1.1, 1.0]
-women_means = [1.8, 1.5, 1.1, 1.3, 0.9]
 
 # Color palette
 light_gray = "#cacaca"
@@ -35,68 +30,95 @@ dark_green = "#33a02c"
 light_red = "#fb9a99"
 dark_red = "#e31a1c"
 
-x = np.arange(len(labels))  # the label locations
-width = 0.35  # the width of the bars
+# Plot an example speedup plot
+def plot_speedup():
+    labels = ['G1', 'G2', 'G3', 'G4', 'G5']
+    men_means = [1.5, 1.2, 1.3, 1.1, 1.0]
+    women_means = [1.8, 1.5, 1.1, 1.3, 0.9]
 
-fig, ax = plt.subplots()
-rects1 = ax.bar(x - width / 2, men_means, width, label='Men', color=light_blue)
-rects2 = ax.bar(x + width / 2,
-                women_means,
-                width,
-                label='Women',
-                color=dark_blue)
+    x = np.arange(len(labels))  # the label locations
+    width = 0.35  # the width of the bars
 
-# Y-Axis Label
-#
-# Use a horizontal label for improved readability.
-ax.set_ylabel('Speedup',
-              rotation='horizontal',
-              position=(1, 1.05),
-              horizontalalignment='left',
-              verticalalignment='bottom')
+    fig, ax = plt.subplots()
+    rects1 = ax.bar(x - width / 2,
+                    men_means,
+                    width,
+                    label='Men',
+                    color=light_blue)
+    rects2 = ax.bar(x + width / 2,
+                    women_means,
+                    width,
+                    label='Women',
+                    color=dark_blue)
 
-# Add some text for labels, title and custom x-axis tick labels, etc.
-ax.set_xticks(x)
-ax.set_xticklabels(labels)
+    # Y-Axis Label
+    #
+    # Use a horizontal label for improved readability.
+    ax.set_ylabel('Speedup',
+                  rotation='horizontal',
[…]

### 2022-11-13 plots/plot.py
plotting: factor out file creation

@@ -30,6 +30,18 @@ dark_green = "#33a02c"
 light_red = "#fb9a99"
 dark_red = "#e31a1c"
 
+
+def save(figure, name):
+    # Do not emit a creation date, creator name, or producer. This will make the
+    # content of the pdfs we generate more deterministic.
+    metadata = {'CreationDate': None, 'Creator': None, 'Producer': None}
+
+    figure.savefig(name, metadata=metadata)
+
+    # Close figure to avoid warning about too many open figures.
+    plt.close(figure)
+
+
 # Plot an example speedup plot
 def plot_speedup():
     labels = ['G1', 'G2', 'G3', 'G4', 'G5']
@@ -96,12 +108,7 @@ def plot_speedup():
 
     fig.tight_layout()
 
-
-    # Do not emit a creation date. This will make the content of
-    # the pdfs we generate more deterministic.
-    metadata = {'CreationDate': None}
-
-    fig.savefig("speedup.pdf", metadata=metadata)
+    save(fig, 'speedup.pdf')
 
 
 def main():

### 2022-11-13 plots/plot.py
plotting: remove unused import

@@ -1,7 +1,6 @@
 #!/usr/bin/env python3
 
 import argparse
-import os
 
 import matplotlib
 import matplotlib.pyplot as plt

### 2022-11-13 plots/plot.py
plotting: move tight_layout into the global defaults

@@ -17,6 +17,11 @@ def setGlobalDefaults():
     matplotlib.rcParams['pdf.fonttype'] = 42
     matplotlib.rcParams['ps.fonttype'] = 42
 
+    ## Enable tight_layout by default
+    #
+    # This ensures the plot has always sufficient space for legends, ...
+    # Without this sometimes parts of the figure would be cut off.
+    matplotlib.rcParams['figure.autolayout'] = True
 
 matplotlib.rcParams['figure.figsize'] = 5, 2
 
@@ -105,8 +110,6 @@ def plot_speedup():
 
     autolabel(rects1)
     autolabel(rects2)
-
-    fig.tight_layout()
 
     save(fig, 'speedup.pdf')
 

### 2022-11-13 plots/plot.py
plotting: factor out legend formatting

@@ -23,6 +23,10 @@ def setGlobalDefaults():
     # Without this sometimes parts of the figure would be cut off.
     matplotlib.rcParams['figure.autolayout'] = True
 
+    ## Legend defaults
+    matplotlib.rcParams['legend.frameon'] = False
+
+
 matplotlib.rcParams['figure.figsize'] = 5, 2
 
 # Color palette
@@ -82,7 +86,6 @@ def plot_speedup():
     ax.set_xticklabels(labels)
 
     ax.legend(ncol=100,
-              frameon=False,
               loc='lower right',
               bbox_to_anchor=(0, 1, 1, 0))
 

### 2023-04-22 tools/create-video.py
Make minted custom lexers work again
minted 2.7 changed its behaviour to not allow for custom command
line options, which broke our custom lexer definitions. Here is
a similar bug reported: https://github.com/gpoore/minted/issues/360.

This commit adds a new minted option that allows custom lexers to work
wiht newer texlive distributions. We also enhance our CI to check that
the paper builds with old and new textlive distributions.

@@ -201,7 +201,9 @@ def createImage(data):
 
     shutil.copyfile("Makefile", dirname + "/Makefile")
     shutil.copyfile(".latexmkrc", dirname + "/.latexmkrc")
-    run(['make', '-C', dirname, f"{base_filename}.pdf"], stdout=DEVNULL, stderr=STDOUT)
+    cmd = ['make', '-C', dirname, f"{base_filename}.pdf"]
+    print(" ".join(cmd))
+    #run(['make', '-C', dirname, f"{base_filename}.pdf"], stdout=STDOUT, stderr=STDOUT)
     createImageOfPaper(dirname + f"/{base_filename}.pdf")
     plotStatistics(commit,  dirname + "/statistics.png", branch, count)
     run(['convert', dirname + f"/{base_filename}.pdf-full.png", '-resize', '3840x2160',

### 2023-04-22 tools/create-video.py
create-video.py: avoid segfault due to mismatched types

@@ -62,10 +62,10 @@ def createImageOfPaper(path, width = 7, pages = 21):
     targetPage = (3840, 2160)
 
     targetPdf = (targetPage[0] / grid[0], targetPage[1] / grid[1])
-    densityWidth = 100 / pdfSize[0] * (targetPage[0] / grid[0]) / 100 * 72
-    densityHeight = 100 / pdfSize[1] * (targetPage[1] / grid[1]) / 100 * 72
+    densityWidth = 100 / float(pdfSize[0]) * (targetPage[0] / grid[0]) / 100 * 72
+    densityHeight = 100 / float(pdfSize[1]) * (targetPage[1] / grid[1]) / 100 * 72
 
-    density = min(densityWidth, densityHeight)
+    density = int(min(densityWidth, densityHeight))
     rows = grid[1]
 
     cmd = ['convert', '-quality', '100', '-density', str(density), path, '-background', 'white', '-alpha', 'remove', '-alpha', 'off', name + "-tmp.png"]

### 2023-04-22 tools/create-video.py
Add option to provide custom make command

@@ -192,6 +192,7 @@ def createImage(data):
     branch = data[2]
     count = data[3]
     base_filename = data[4]
+    make = data[5]
     filename = str(commit_number) + ".zip"
     dirname =  str(commit_number).zfill(5) + "-files"
 
@@ -201,7 +202,8 @@ def createImage(data):
 
     shutil.copyfile("Makefile", dirname + "/Makefile")
     shutil.copyfile(".latexmkrc", dirname + "/.latexmkrc")
-    run(['make', '-C', dirname, f"{base_filename}.pdf"], stdout=DEVNULL, stderr=STDOUT)
+    cmd = f"{make} -C {dirname} {base_filename}.pdf"
+    run(cmd, shell=True, stdout=DEVNULL, stderr=STDOUT)
     createImageOfPaper(dirname + f"/{base_filename}.pdf")
     plotStatistics(commit,  dirname + "/statistics.png", branch, count)
     run(['convert', dirname + f"/{base_filename}.pdf-full.png", '-resize', '3840x2160',
@@ -211,11 +213,11 @@ def createImage(data):
         dirname + '/statistics.png', '-composite', dirname +
         f"/{base_filename}.pdf-overlay.png"])
 
-def createImages(branch, count, filename, procs):
+def createImages(branch, count, filename, procs, make):
     repo = Repo(".")
     commits = list(reversed(list(repo.iter_commits(branch, max_count=count))))
     commits = list(map(lambda x: x.hexsha, commits))
-    items = list(zip(range(len(commits)), commits, [branch] * len(commits), [count] * len(commits), [filename] * len(commits)))
+    items = list(zip(range(len(commits)), commits, [branch] * len(commits), [count] * len(commits), [filename] * len(commits), [make] * len(commits)))
 
     p = Pool(procs)
     for _ in tqdm.tqdm(p.imap_unordered(createImage, items), total=len(items)):
@@ -258,9 +260,10 @@ if __name__ == "__main__":
     parser.add_argument('-p', '--procs', default=16)
     parser.add_argument('-b', '--branch', default='main')
     parser.add_argument('-f', '--filename', default='paper')
+    parser.add_argument('-m', '--make', default='make')
 
     args = parser.parse_args()
 
 
-    createImages(args.branch, args.count, args.filename, int(args.procs))
+    createImages(args.branch, args.count, args.filename, int(args.procs), args.make)
     createVideo(basename = args.filename)

### 2023-04-22 tools/create-video.py
Provide error message when we cannot build a commit

@@ -205,8 +205,15 @@ def createImage(data):
     cmd = f"{make} -C {dirname} {base_filename}.pdf"
     run(cmd, shell=True, stdout=DEVNULL, stderr=STDOUT)
     createImageOfPaper(dirname + f"/{base_filename}.pdf")
+
+    fullpng = dirname + f"/{base_filename}.pdf-full.png"
+
+    if not os.path.exists(fullpng):
+        print(f"Could not build commit {commit_number} with hash {commit}.")
+        return
+
     plotStatistics(commit,  dirname + "/statistics.png", branch, count)
-    run(['convert', dirname + f"/{base_filename}.pdf-full.png", '-resize', '3840x2160',
+    run(['convert', fullpng, '-resize', '3840x2160',
         '-background', 'white', '-gravity', 'center', '-extent', '3840x2160',
         dirname + f"/{base_filename}.pdf-expanded.png"])
     run(['convert', '-gravity', 'SouthEast', dirname + f"/{base_filename}.pdf-expanded.png",

### 2023-12-30 paper.tex
add first draft

--- 2021-04-21 paper.tex
@@
 \def\paperversiondraft{draft}
-\def\paperversionblind{blind}
-\def\paperversioncamera{camera}
+\def\paperversionblind{normal}
+\def\paperversioncameraACM{cameraACM}
 
-% If no special paper-version is requested, compile in draft mode
-\ifx\paperversion\paperversionblind
+% If no draft paper-version is requested, compile in 'normal' mode
+\ifx\paperversion\paperversiondraft
 \else
-  \ifx\paperversion\paperversioncamera
+  \ifx\paperversion\paperversioncameraACM
   \else
-     \def\paperversion{draft}
+  \def\paperversion{normal}
   \fi
@@
 
-\def\grammarlyon{on}
-
-\ifx\grammarly\grammarlyon
-\def\review{}
+\ifx\paperversion\paperversioncameraACM
+  \documentclass[sigplan]{acmart}
 \else
-\def\review{review,}
+  \documentclass[review, anonymous, sigplan]{acmart}
 \fi
 
-\ifx\paperversion\paperversiondraft
-  \documentclass[11pt, manuscript,\review anonymous]{acmart}
+\def\acmversionanonymous{anonymous}
+\def\acmversionjournal{journal}
+\def\acmversionnone{none}
+
+\makeatletter
+\if@ACM@anonymous
+  \def\acmversion{anonymous}
+\else
+  \def\acmversion{journal}
 \fi
+\makeatother
 
-\ifx\paperversion\paperversionblind
-  \documentclass[manuscript,\review anonymous]{acmart}
-\fi
-
-\ifx\paperversion\paperversioncamera
-  \documentclass[11pt, manuscript, nonacm, natbib, oneside]{acmart}\settopmatter{}
-\fi
-
-\geometry{twoside=false}
 \usepackage{colortbl}
@@
 \usepackage{lipsum}
-\usepackage[textsize=tiny]{todonotes}
 \usepackage{xparse}
 \usepackage{xifthen, xstring}
-\usepackage[normalem]{ulem}
 \usepackage{xspace}
 \usepackage{marginnote}
+\usepackage{etoolbox}
+
+\input{tex/setup.tex}
+\input{tex/acm.tex}
+
+\usemintedstyle{colorful}
+
+% Newer versions of minted require the 'customlexer' argument for custom lexers
+% whereas older versions require the '-x' to be passed via the command line.
 \makeatletter
-\font\uwavefont=lasyb10 scaled 652
-
-\newcommand\colorwave[1][blue]{\bgroup\markoverwith{\lower3\p@\hbox{\uwavefont\textcolor{#1}{\char58}}}\ULon}
-\makeatother
-
-\ifx\paperversion\paperversiondraft
-\newcommand\createtodoauthor[2]{%
-\def\tmpdefault{emptystring}
-\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
-   {\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}}
-   {\ifthenelse{\equal{##2}{}}{\colorwave[#2]{##1}\xspace}{\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}\colorwave[#2]{##1}}}}
-\expandafter\newcommand\csname #1f\endcsname[2][\tmpdefault]{
-	\marginnote{
-		\todo[inline,linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1 (Figure):} ##2}}
+\ifcsdef{minted@optlistcl@quote}
+{
+\newminted[mlir]{tools/MLIRLexer.py:MLIRLexerOnlyOps}{customlexer, mathescape}
+\newminted[xdsl]{tools/MLIRLexer.py:MLIRLexer}{customlexer, mathescape, style=murphy}
+\newminted[lean4]{tools/Lean4Lexer.py:Lean4Lexer}{customlexer, mathescape}
 }
-}
-
-\else
-\newcommand\createtodoauthor[2]{%
-\def\tmpdefault{emptystring}
-\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
-   {}
-   {\ifthenelse{\equal{##2}{}}{##1\xspace}{##1}}}
-\expandafter\newcommand\csname #1f\endcsname[2][\tmpdefault]{
-   }
-   }
-\fi
-
-% Broaden margins to make room for todo notes
-\makeatletter
-\patchcmd{\@addmarginpar}{\ifodd\c@page}{\ifodd\c@page\@tempcnta\m@ne}{}{}
-\makeatother
-\ifx\paperversion\paperversiondraft
-  \makeatletter
-  \if@ACM@journal
-    \geometry{asymmetric}
-    \paperwidth=\dimexpr \paperwidth + 3.5cm\relax
-    \oddsidemargin=\dimexpr\oddsidemargin + 0cm\relax
-    \evensidemargin=\dimexpr\evensidemargin + 0cm\relax
-    \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
-    \setlength{\marginparwidth}{4.6cm}
-    % This makeatletter box helps to move notes to the right
-    \makeatletter
-    \long\def\@mn@@@marginnote[#1]#2[#3]{%
-      \begingroup
-        \ifmmode\mn@strut\let\@tempa\mn@vadjust\else
-          \if@inlabel\leavevmode\fi
-          \ifhmode\mn@strut\let\@tempa\mn@vadjust\else\let\@tempa\mn@vlap\fi
-        \fi
-        \@tempa{%
-          \vbox to\z@{%
-            \vss
-            \@mn@margintest
-            \if@reversemargin\if@tempswa
-                \@tempswafalse
-              \else
-                \@tempswatrue
-            \fi\fi
-            %\if@tempswa
-              \rlap{%
-                \if@mn@verbose
-                  \PackageInfo{marginnote}{xpos seems to be \@mn@currxpos}%
-                \fi
-                \begingroup
-                  \ifx\@mn@currxpos\relax\else\ifx\@mn@currxpos\@empty\else
-                      \kern-\dimexpr\@mn@currxpos\relax
-                  \fi\fi
-                  \ifx\@mn@currpage\relax
-                    \let\@mn@currpage\@ne
-                  \fi
-                  \if@twoside\ifodd\@mn@currpage\relax
-                      \kern\oddsidemargin
-                    \else
-                      \kern\evensidemargin
-                    \fi
-                  \else
-                    \kern\oddsidemargin
-                  \fi
-                  \kern 1in
-                \endgroup
-                \kern\marginnotetextwidth\kern\marginparsep
-                \vbox to\z@{\kern\marginnotevadjust\kern #3
-                  \vbox to\z@{%
-                    \hsize\marginparwidth
-                    \linewidth\hsize
-                    \kern-\parskip
-                    \marginfont\raggedrightmarginnote\strut\hspace{\z@}%
-                    \ignorespaces#2\endgraf
-                    \vss}%
-                  \vss}%
-              }%
-          }%
-        }%
-      \endgroup
+{
+\newminted[mlir]{tools/MLIRLexer.py:MLIRLexerOnlyOps -x}{mathescape}
+\newminted[xdsl]{tools/MLIRLexer.py:MLIRLexer -x}{mathescape, style=murphy}
+\newminted[lean4]{tools/Lean4Lexer.py:Lean4Lexer -x}{mathescape}
 }
 \makeatother
-  \else
-    \paperwidth=\dimexpr \paperwidth + 6cm\relax
-    \oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
-    \evensidemargin=\dimexpr\evensidemargin + 3cm\relax
-    \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
-    \setlength{\marginparwidth}{4.6cm}
-  \fi
-  \makeatother
-\fi
 
@@
 
-% Print \autoref as "Section X.Y.Z"
-\renewcommand*{\sectionautorefname}{Section}
-\renewcommand*{\subsectionautorefname}{Section}
-\renewcommand*{\subsubsectionautorefname}{Section}
+\usepackage{booktabs}
+\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
+
+\usepackage[verbose]{newunicodechar}
+\newunicodechar{₁}{\ensuremath{_1}}
+\newunicodechar{₂}{\ensuremath{_2}}
+\newunicodechar{∀}{\ensuremath{\forall}}
+\newunicodechar{α}{\ensuremath{\alpha}}
+\newunicodechar{β}{\ensuremath{\beta}}
 
@@
 \usetikzlibrary{shapes}
-\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
-            \node[shape=circle,fill=pairedOneLightBlue,inner sep=1pt] (char) {#1};}}
+\DeclareRobustCommand\circled[2][]{\ifthenelse{\isempty{#1}}{\tikz[baseline=(char.base)]{\node[shape=circle,fill=pairedOneLightBlue,inner sep=1pt] (char) {#2};}}{\autoref{#1}: \hyperref[#1]{\tikz[baseline=(char.base)]{\node[shape=circle,fill=pairedOneLightBlue,inner sep=1pt] (char) {#2};}}}}
 
-%% Note: Authors migrating a paper from traditional SIGPLAN
-%% proceedings format to PACMPL format should change 'sigplan' to
-%% 'acmsmall'.
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-\usepackage{bibentry}
-\nobibliography*
-
-\makeatletter\if@ACM@journal\makeatother
-%% Journal information (used by PACMPL format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{1}
-\acmArticle{1}
-\acmYear{2017}
-\acmMonth{1}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\else\makeatother
-%% Conference information (used by SIGPLAN proceedings format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
-\acmYear{2017}
-\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\fi
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission
-\setcopyright{none}             %% For review submission
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2017}           %% If different from \acmYear
-
-%% Bibliography style
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-%\citestyle{acmauthoryear}  %% For author/year citations
-%\citestyle{acmnumeric}     %% For numeric citations
-%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
-                            %% sorting of references within a single citation;
-                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
-                            %% rendered as [14,5,2] rather than [2,5,14].
-%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
-                            %% compression of sequential references within a
-                            %% single citation;
-                            %% e.g., \cite{Baker12,Baker14,Baker16}
-                            %% rendered as [2,3,4] rather than [2-4].
+% listings don't write "Listing" in autoref without this.
+\providecommand*{\listingautorefname}{Listing}
 
@@
 %% Title information
-\title[]{
-	{
-	\centering
-	Scientific Report - Summary of Research\\
-	~\\
-Automatized compilation of sequential software to a diverse set of
-hardware accelerators\\
-}
-~\\
-\vspace{18em}
-\large
-Grant number: PZ00P2 168016\\
-Period: 01.05.2018– 31.12.2020\\
-Grantee: Tobias Grosser\\
-\today
-}         %% [Short Title] is optional;
+\title{Fast \& Correct Performance Programming in Lean4}       %% [Short Title] is optional;
                                       %% when present, will be used in
                                       %% header instead of Full Title.
-
 %% Author information
@@
 %% extraction tools.
+\author{Siddharth Bhat}
+\authornote{with author1 note}          %% \authornote is optional;
+                                      %% can be repeated if necessary
+\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position1}
+  \department{Department1}              %% \department is recommended
+  \institution{Institution1}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{City1}
+  \state{State1}
+  \postcode{Post-Code1}
+  \country{Country1}
+}
+\email{first1.last1@inst1.edu}          %% \email is recommended
 
-%% Author with single affiliation.
+\author{Henrik Boving}
+\authornote{with author2 note}          %% \authornote is optional;
+                                      %% can be repeated if necessary
+\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position2a}
+  \department{Department2a}             %% \department is recommended
+  \institution{Institution2a}           %% \institution is required
+  \streetaddress{Street2a Address2a}
+  \city{City2a}
+  \state{State2a}
+  \postcode{Post-Code2a}
+  \country{Country2a}
+}
+\email{first2.last2@inst2a.com}         %% \email is recommended
+\affiliation{
+  \position{Position2b}
+  \department{Department2b}             %% \department is recommended
+  \institution{Institution2b}           %% \institution is required
+  \streetaddress{Street3b Address2b}
+  \city{City2b}
+  \state{State2b}
+  \postcode{Post-Code2b}
+  \country{Country2b}
+}
+\email{first2.last2@inst2b.org}         %% \email is recommended
 
-% An abstract should consist of six main sentences:
-%  1. Introduction. In one sentence, what’s the topic?
-%  2. State the problem you tackle.
-%  3. Summarize (in one sentence) why nobody else has adequately answered the research question yet.
-%  4. Explain, in one sentence, how you tackled the research question.
-%  5. In one sentence, how did you go about doing the research that follows from your big idea.
-%  6. As a single sentence, what’s the key impact of your research?
+\author{Tobias Grosser}
+\authornote{with author2 note}          %% \authornote is optional;
+                                      %% can be repeated if necessary
+\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position2a}
+  \department{Department2a}             %% \department is recommended
+  \institution{Institution2a}           %% \institution is required
+  \streetaddress{Street2a Address2a}
+  \city{City2a}
+  \state{State2a}
+  \postcode{Post-Code2a}
+  \country{Country2a}
+}
+\email{first2.last2@inst2a.com}         %% \email is recommended
+\affiliation{
+  \position{Position2b}
+  \department{Department2b}             %% \department is recommended
+  \institution{Institution2b}           %% \institution is required
+  \streetaddress{Street3b Address2b}
+  \city{City2b}
+  \state{State2b}
+  \postcode{Post-Code2b}
+  \country{Country2b}
+}
+\email{first2.last2@inst2b.org}         %% \email is recommended
 
-% (http://www.easterbrook.ca/steve/2010/01/how-to-write-a-scientific-abstract-in-six-easy-steps/)
+\begin{abstract}
+
+ Software bugs and security vulnerabilities pose enormous costs and risks in today’s highly interconnected world.
+ Exceptionally critical programs are today already written together with formal proofs of their correctness.
+ However, code that is easy to reason about and to prove correct often runs unacceptably slowly, while code that runs fast is very hard to reason about.
+ As a result, formal proofs are not used for performance-critical code.
+ With the increased use of high-performance computations in critical cyber-physical systems (e.g., deep neural networks for autonomous driving),
+ having to choose between fast and provably correct software is not acceptable any more.
+ By designing novel techniques for the development of trustworthy high-performance software, we will bridge this gap and eliminate the need to trade-off between fast and correct.
+
+We plan to present a platform for building such high assurance and performance programming,
+by retooling the Lean 4 proof assistant for verified code generation.
+We have already deeply embedded support for performant code generation by embedding the LLVM backend in Lean4.
+We are currently working towards exposing the LLVM backend to be user-facing. 
+This enables end-users to control the code generation in a proof assistant more finely than has been possible before.
+We believe that our work sets the foundation to enable programmers to write programs that are both fast and correctness.
+
+\end{abstract}
 
@@
 % Drop citations and footnotes in draft and blind mode.
-\ifx\paperversion\paperversioncamera
+\ifx\acmversion\acmversionanonymous
+\settopmatter{printacmref=false} % Removes citation information below abstract
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+\fi
+\ifx\acmversion\acmversionjournal
 %% 2012 ACM Computing Classification System (CSS) concepts
 %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011008</concept_id>
+<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003456.10003457.10003521.10003525</concept_id>
+<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~General programming languages}
+\ccsdesc[300]{Social and professional topics~History of programming languages}
 %% End of generated code
@@
 %% comma separated list
-
-\settopmatter{printacmref=false} % Removes citation information below abstract
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+\keywords{keyword1, keyword2, keyword3} 
 \fi
@@
 \maketitle
-\newpage
-\ifx\grammarly\grammarlyon 
-\onecolumn 
-\else 
-\fi
-
-\section{Research Work Conducted}
-
-The primary goals of this Ambizione research project were proposed as:
-\begin{itemize}
-\item Develop a multi-target heterogeneous compute compiler for C++ and Julia code that generates binaries which effectively exploit
-OpenCL-capable accelerators including GPU, Xeon-Phi, and FPGA.
-\item Research new techniques to scale polyhedral optimization to large kernels and to improve their efficiency to a level that permits
-the effective exploration of a large set of mapping choices.
-\item Derive effective performance models based on hybrid analytical/machine-learning techniques and use them to analyze and optimize accelerator mapping decisions in a large body of existing software.
-\item Evolve our new performance modeling techniques to work within the resource constraints of Julias dynamic run-time system and use Julia to experiment with distributed learning techniques.
-\end{itemize}
-
-The focus of year three in this Ambizione project was the integration of a scalable
-polyhedral core infrastructure into real-world compilers (e.g, LLVM/MLIR),
-establishing backends for targetting accelerators such as GPUs and FPGAs, and,
-finally, developing good performance models. As mentioned in the year 2
-report, our focus moved from Julia to MLIR as a compiler and programming
-language infrastructure.
-
-In the following, we discuss four milestones we identified as important to
-reach these goals:
-
-\begin{enumerate}
-\item Fast, robust, and correct polyhedral libraries
-\item Automatic Raising to high-level descriptions
-\item Fast GPU code generation demonstrated via a climate DSL
-\item Code generation for emerging architectures and FPGA
-\item Scalable performance analysis within a compiler.
-\end{enumerate}
-
-\subsection{Fast, robust, and correct polyhedral libraries}
-Presburger arithmetic provides the mathematical core for the polyhedral
-compilation techniques that drive analytical cache models, loop optimization
-for ML and HPC, formal verification, and even hardware design.  Polyhedral
-compilation is widely regarded as being slow due to the potentially high
-computational cost of the underlying Presburger libraries. Researchers
-typically use these libraries as powerful black-box tools, but a lack of
-internal documentation and decade-old C implementations hold back broader
-performance-optimization efforts. With FPL, we introduced a new library for
-Presburger arithmetic for the MLIR compiler built from the ground up in modern
-C++. We carefully documented its internal algorithmic foundations, use
-lightweight C++ data structures to minimize memory management costs, and deployed
-transprecision computing across the entire library to effectively exploit
-machine integers and vector instructions. On a newly-developed comprehensive
-benchmark suite for Presburger arithmetic, we show a 5.25x speedup in total
-runtime over the state-of-the-art library isl in its default configuration and
-3.14x when over a variant of isl optimized with element-wise transprecision
-computing. We expect that the availability of a well-documented and fast
-Presburger library will accelerate the adoption of polyhedral compilation
-techniques in production compilers.
-
-The central components of this library have already been upstreamed into the
-open-source LLVM/MLIR project and are usable by the wider research and 
-industrial community. The work in this subproject has been published in \citet{grosser2020fast} and
-is under review in \citet{pichanatan2021}.
-
-\subsection{Automatic Raising to high-level descriptions}
-
-Multi-level intermediate representations (IR) show great promise for lowering
-the design costs for domain-specific compilers by providing a reusable,
-extensible, and non-opinionated framework for expressing domain-specific and
-high-level abstractions directly in the IR. But, while such frameworks support
-the progressive lowering of high-level representations to low-level IR, they do
-not raise in the opposite direction. Thus, the entry point into the compilation
-pipeline defines the highest level of abstraction for all subsequent
-transformations, limiting the set of applicable optimizations, in particular
-for general-purpose languages that are not semantically rich enough to model
-the required abstractions.  We propose Progressive Raising, a complementary
-approach to the progressive lowering in multi-level IRs that raises from lower
-to higher-level abstractions to leverage domain-specific transformations for
-low-level representations. We further introduce MultiLevel Tactics, our
-declarative approach for progressive raising, implemented on top of the MLIR
-framework, and demonstrate the progressive raising from affine loop nests
-specified in a general-purpose language to high-level linear algebra
-operations.  Our raising paths leverage subsequent high-level domain-specific
-transformations with significant performance improvements.
-
-All this work has been has already been upstreamed into the open-source
-LLVM/MLIR project and is usable by the wider research and industrial community.
-The work in this subproject has been published in \citet{chelini2021mlir}.
-
-\subsection{Fast GPU code generation demonstrated via a climate DSL}
-
-Most compilers have a single core intermediate representation (IR) (e.g., LLVM)
-sometimes complemented with vaguely-defined IR-like data structures. This IR is
-commonly low-level and close to machine instructions. As a result,
-optimizations relying on domain-specific information are either not possible or
-require complex analysis to recover the missing information. In contrast,
-multi-level rewriting instantiates a hierarchy of dialects (IRs), lowers
-programs level-by-level, and performs code transformations at the most suitable
-level. We demonstrate the effectiveness of this approach for the weather and
-climate domain. In particular, we develop a prototype compiler and design
-stencil- and GPU-specific dialects based on a set of newly introduced design
-principles. We find that two domain-specific optimizations (500 lines of code)
-realized on top of LLVM’s extensible MLIR compiler infrastructure suffice to
-outperform state-of-the-art solutions. In essence, multi-level rewriting
-promises to herald the age of specialized compilers composed from domain- and
-target-specific dialects implemented on top of a shared infrastructure.
-
-The GPU backend of the open-earth compiler has already been upstreamed into
-the open-source LLVM/MLIR project and is useable by the wider research
-and industrial community. The work in this subproject has been published in \citet{gysi2020ooc}
-
-\subsection{Code Generation for Emerging Architectures and FPGA}
-
-Automatic optimization for emerging architectures and FPGA has been a principal
-objective of this work. We exploited the polyhedral techniques we developed in
-earlier tasks to generate high-performance code for emerging memory
-architectures such as racetrack memories to optimize polyhedral loop nests and
-deep neural networks. We also proposed infrastructure for an open-source
-hardware stack that allows for the compilation and simulation of System Verilog
-code. We also observed an increased use of systems with where the data models
-(e.g., pointer sizes) differ between host systems and accelerators. To ensure
-the automatic code generation for such architectures that also allows for
-automatic data movement, we expanded LLVM and its OpenMP software stack to
-support mixed-data-model offloading.
-
-The LLHD compiler has been upstreamed as open-source into the LLVM/CIRCT
-project. The work in this subproject has been published in \citet{khan2020polyhedral},
-\citet{schueki2020llhd}, \citet{kurth2020mixed}, and \citet{kourtis2020compiling}.
-
-\subsection{Scalable performance analysis within a compiler}
-
-Performance models are well-known instruments to understand the scaling
-behavior of parallel applications. They express how performance changes as key
-execution parameters, such as the number of processes or the size of the input
-problem, vary. Besides reasoning about program behavior, such models can also
-be automatically derived from performance data. This is called empirical
-performance modeling. While this sounds simple at the first glance, this
-approach faces several serious interrelated challenges, including expensive
-performance measurements, inaccuracies inflicted by noisy benchmark data, and
-overall complex experiment design, starting with the selection of the right
-parameters. The more parameters one considers, the more experiments are needed
-and the stronger the impact of noise. In this paper, we show how taint
-analysis, a technique borrowed from the domain of computer security, can
-substantially improve the modeling process, lowering its cost, improving model
-quality, and help validate performance models and experimental setups.
-
-In a subsequent step, we analyzed the optimality of function inlining as a means
-to minimize code size and alter expose further optimization opportunities. This
-work has shown promising code size reductions on SPEC 2017 and the LLVM compiler.
-We hope to use it in the future as an enabling transformation for large-scale
-automatic accelerator mapping.
-
-The work in this subproject has been published in \citet{copic2021} and is under review in
-\citet{theo2021}.
-
-\section{Deviations}
-
-\section{Important Events}
-
-\section{Contributions}
-
-\section{Research Output}
-
-\subsection{Publications (published)}
-\begin{enumerate}
-	\item \bibentry{chelini2021mlir}
-	\item \bibentry{10.1145/3437801.3441613}
-	\item \bibentry{grosser2020fast}
-	\item \bibentry{chelini2020automatic}
-	\item \bibentry{khan2020polyhedral}
-	\item \bibentry{schueki2020llhd}
-	\item \bibentry{kurth2020mixed}
-	\item \bibentry{kourtis2020compiling}
-\end{enumerate}
-\subsection{Publications (under review)}
-\begin{enumerate}
-	\item \bibentry{gysi2020ooc}
-	\item \bibentry{fehr2021}
-	\item \bibentry{anurudth2021}
-	\item \bibentry{theo2021}
-	\item \bibentry{pichanatan2021}
-	\item \bibentry{copic2021}
-\end{enumerate}
-
-%% Bibliography
-\bibliographystyle{plainnat}
-\nobibliography{references}
 

### 2024-09-24 plots/plot.py
Add information

@@ -6,8 +6,10 @@ import matplotlib
 import matplotlib.pyplot as plt
 import numpy as np
 import math
+import csv
 
 from typing import Callable
+
 
 def setGlobalDefaults():
     ## Use TrueType fonts instead of Type 3 fonts
@@ -26,7 +28,7 @@ def setGlobalDefaults():
 
     ## Legend defaults
     matplotlib.rcParams['legend.frameon'] = False
-    
+
     # Hide the right and top spines
     #
     # This reduces the number of lines in the plot. Lines typically catch
@@ -36,7 +38,6 @@ def setGlobalDefaults():
     matplotlib.rcParams['axes.spines.right'] = False
     matplotlib.rcParams['axes.spines.top'] = False
 
-matplotlib.rcParams['figure.figsize'] = 5, 2
 
 # Color palette
 light_gray = "#cacaca"
@@ -50,6 +51,7 @@ dark_red = "#e31a1c"
 black = "#000000"
 white = "#ffffff"
 
+
 def save(figure, name):
     # Do not emit a creation date, creator name, or producer. This will make the
     # content of the pdfs we generate more deterministic.
@@ -59,8 +61,9 @@ def save(figure, name):
 
     # Close figure to avoid warning about too many open figures.
     plt.close(figure)
-    
+
     print(f'written to {name}')
+
 
 # helper for str_from_float.
 # format float in scientific with at most *digits* digits.
@@ -71,18 +74,19 @@ def save(figure, name):
 def get_scientific(x: float, digits: int):
     # get scientific without leading zeros or + in exp
     def get(x: float, prec: int) -> str:
-      result = f'{x:.{prec}e}'
-      result = result.replace('e+', 'e')
-      while 'e0' in result:
-        result = result.replace('e0', 'e')
-      while 'e-0' in result:
-        result = result.replace('e-0', 'e-')
-      return result
+        result = f'{x:.{prec}e}'
+        result = result.replace('e+', 'e')
+        while 'e0' in result:
+            result = result.replace('e0', 'e')
+        while 'e-0' in result:
+            result = result.replace('e-0', 'e-')
+        return result
 
     result = get(x, digits)
     len_after_e = len(result.split('e')[1])
     prec = max(0, digits - len_after_e - 2)
     return get(x, prec)
+
 
 # format float with at most *digits* digits.
 # if the number is too small or too big,
[…]

### 2025-02-03 tools/lexers/MLIRLexer.py
Update MLIR lexer

@@ -5,20 +5,35 @@ comment_rule = (r'//.*?\n', Comment)
 ssa_value_rule = (r'%[^[ )]*]*', Name.Variable)
 symbol_rule = (r'@[^(]*', Name.Function)
 basic_block_rule = (r'\^[^(:\]]*', Name.Label)
-operation_rule = (r'(=)( +)([a-z_]+)(\.)([a-z_]+)',
-                  bygroups(Operator, Text, Name.Namespace, Text,
-                           Keyword.Function))
-non_assign_operation_rule = (r'([a-z_]+)(\.)([a-z_]+)',
-                             bygroups(Name.Namespace, Text, Keyword.Function))
-type_rule = (r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
-             bygroups(Operator, Name.Namespace, Text, Keyword.Type))
+operation_rule = (
+    r'(=)( +)([a-z_]+)(\.)([a-z_\.]+)',
+    bygroups(Operator, Text, Name.Namespace, Text, Keyword.Function),
+)
+non_assign_operation_rule = (
+    r'([a-z_]+)(\.)([a-z_\.]+)',
+    bygroups(Name.Namespace, Text, Keyword.Function),
+)
+builtin_type_rule = (
+    r'(#)([a-z_]+)(\.)([a-z0-9_]+)',
+    bygroups(Operator, Name.Namespace, Text, Keyword.Type),
+)
+type_rule = (
+    r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
+    bygroups(Operator, Name.Namespace, Text, Keyword.Type),
+)
 abbrev_type_tule = (r'(!)([a-z0-9]+)', bygroups(Operator, Keyword.Type))
-first_attribute_rule = (r'([{\[])([a-z_A-Z]+)( = +)([@a-z0-9">=]+)',
-                        bygroups(Text, Name.Attribute, Text, Name.Tag))
-following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=@]+)',
-                            bygroups(Text, Name.Attribute, Text, Name.Tag))
-abbrev_following_attribute_rule = (r'(, +)([a-z_]+)( = +)',
-                                   bygroups(Text, Name.Attribute, Text))
+first_attribute_rule = (
+    r'([{\[])([a-z_A-Z]+)( = +)([@a-z0-9">=]+)',
[…]

### 2025-02-03 tools/lexers/MLIRLexer.py
Add number literals to lexer

@@ -1,5 +1,5 @@
 from pygments.lexer import RegexLexer, bygroups
-from pygments.token import Name, Keyword, Operator, Comment, Text, Punctuation, Literal
+from pygments.token import Name, Keyword, Number, Operator, Comment, Text, Punctuation, Literal
 
 comment_rule = (r'//.*?\n', Comment)
 ssa_value_rule = (r'%[^[ )]*]*', Name.Variable)
@@ -51,6 +51,7 @@ class MLIRLexer(RegexLexer):
             non_assign_operation_rule,
             builtin_type_rule,
             type_rule,
+            (r'(-?[0-9]+(?:\.[0-9]+)?)', Number),
             abbrev_type_tule,
             (r'(\n|\s)+', Text),
             first_attribute_rule,

### 2025-02-03 tools/lexers/MLIRLexer.py
Update MLIR lexer (#63)
* Update MLIR lexer

* Update config

* Add number literals to lexer

---------

Co-authored-by: Alex Rice <alexrice999@hotmail.co.uk>

@@ -1,24 +1,39 @@
 from pygments.lexer import RegexLexer, bygroups
-from pygments.token import Name, Keyword, Operator, Comment, Text, Punctuation, Literal
+from pygments.token import Name, Keyword, Number, Operator, Comment, Text, Punctuation, Literal
 
 comment_rule = (r'//.*?\n', Comment)
 ssa_value_rule = (r'%[^[ )]*]*', Name.Variable)
 symbol_rule = (r'@[^(]*', Name.Function)
 basic_block_rule = (r'\^[^(:\]]*', Name.Label)
-operation_rule = (r'(=)( +)([a-z_]+)(\.)([a-z_]+)',
-                  bygroups(Operator, Text, Name.Namespace, Text,
-                           Keyword.Function))
-non_assign_operation_rule = (r'([a-z_]+)(\.)([a-z_]+)',
-                             bygroups(Name.Namespace, Text, Keyword.Function))
-type_rule = (r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
-             bygroups(Operator, Name.Namespace, Text, Keyword.Type))
+operation_rule = (
+    r'(=)( +)([a-z_]+)(\.)([a-z_\.]+)',
+    bygroups(Operator, Text, Name.Namespace, Text, Keyword.Function),
+)
+non_assign_operation_rule = (
+    r'([a-z_]+)(\.)([a-z_\.]+)',
+    bygroups(Name.Namespace, Text, Keyword.Function),
+)
+builtin_type_rule = (
+    r'(#)([a-z_]+)(\.)([a-z0-9_]+)',
+    bygroups(Operator, Name.Namespace, Text, Keyword.Type),
+)
+type_rule = (
+    r'(!)([a-z_]+)(\.)([a-z0-9_]+)',
+    bygroups(Operator, Name.Namespace, Text, Keyword.Type),
+)
 abbrev_type_tule = (r'(!)([a-z0-9]+)', bygroups(Operator, Keyword.Type))
-first_attribute_rule = (r'([{\[])([a-z_A-Z]+)( = +)([@a-z0-9">=]+)',
-                        bygroups(Text, Name.Attribute, Text, Name.Tag))
-following_attribute_rule = (r'(, +)([a-z_]+)( = +)([a-z0-9">=@]+)',
-                            bygroups(Text, Name.Attribute, Text, Name.Tag))
-abbrev_following_attribute_rule = (r'(, +)([a-z_]+)( = +)',
-                                   bygroups(Text, Name.Attribute, Text))
[…]

### 2025-06-23 tools/check_latexminted_config_exists.sh
Add automatic generation of .latexminted_config to Makefile

@@ -1,0 +1,20 @@
+#!/usr/bin/env bash
+
+FILE='.latexminted_config'
+
+if [[ !( -f "$HOME/$FILE" || -f "$TEXMF/$FILE" ) ]]; then
+    echo 'Error:'
+    echo "  Neither $HOME/$FILE (~/$FILE) nor $TEXMF/$FILE (\$TEXMF/$FILE) seem to exist!"
+    echo 
+    echo '  To create the file, please run: '
+    echo '    ./tools/create_latexminted_config.sh'  
+    echo ''
+    echo '  Or, manually create the file, with at least the following options set:'
+    echo
+    echo '  {'
+    echo '    "security": {' 
+    echo '      "enable_cwd_config": true' 
+    echo '    }' 
+    echo '  }' 
+    exit 1
+fi

### 2025-06-23 tools/check_latexminted_config_exists.sh
add whitespace to satisfy shellcheck

@@ -2,7 +2,7 @@
 
 FILE='.latexminted_config'
 
-if [[ !( -f "$HOME/$FILE" || -f "$TEXMF/$FILE" ) ]]; then
+if [[ ! ( -f "$HOME/$FILE" || -f "$TEXMF/$FILE" ) ]]; then
     echo 'Error:'
     echo "  Neither $HOME/$FILE (~/$FILE) nor $TEXMF/$FILE (\$TEXMF/$FILE) seem to exist!"
     echo 

### 2025-06-23 tools/check_latexminted_config_exists.sh
Add automatic generation of .latexminted_config to Makefile (#69)
This incorporates the setup notes regarding minted 3's configuration file into the Makefile, so that no more manual setup steps are required.

In particular:
* We add a Makefile recipe for the .latexminted_config file in the current directory, and
* We check for the existence of `~/.latexminted_config` and raise an error if it doesn't
   *  The error points to a new script which can be used to auto-generate this config, too, but I didn't want to pollute the home-dir without explicit prompting from the user
   * Note that we only

@@ -1,0 +1,20 @@
+#!/usr/bin/env bash
+
+FILE='.latexminted_config'
+
+if [[ ! ( -f "$HOME/$FILE" || -f "$TEXMF/$FILE" ) ]]; then
+    echo 'Error:'
+    echo "  Neither $HOME/$FILE (~/$FILE) nor $TEXMF/$FILE (\$TEXMF/$FILE) seem to exist!"
+    echo 
+    echo '  To create the file, please run: '
+    echo '    ./tools/create_latexminted_config.sh'  
+    echo ''
+    echo '  Or, manually create the file, with at least the following options set:'
+    echo
+    echo '  {'
+    echo '    "security": {' 
+    echo '      "enable_cwd_config": true' 
+    echo '    }' 
+    echo '  }' 
+    exit 1
+fi

### 2025-06-23 tools/create_latexminted_config.sh

@@ -1,0 +1,13 @@
+#!/usr/bin/env bash
+
+FILE="${HOME}/.latexminted_config"
+
+if [[ -f "$FILE" ]]; then
+    echo "Error: ${FILE} already exists"
+fi
+
+echo '{
+  "security": {
+      "enable_cwd_config": true
+  }
+}' > "$FILE"

### 2025-06-23 tools/generate_lexers_json.py

@@ -1,3 +1,5 @@
+#!/usr/bin/env python3
+
 import os
 import hashlib
 import json

### 2025-06-24 tools/generate_lexers_json.py
Preserve customized latexminted configuration when refreshing lexer SHAs (#70)
This PR modifies the `generate_lexers_json` script to preserve
configuration in the .latexminted_config file outside of the
custom_lexers option, when refreshing the SHAs.

@@ -3,6 +3,7 @@
 import os
 import hashlib
 import json
+from pathlib import Path
 
 LEXER_DIR = "tools/lexers"
 
@@ -27,13 +28,25 @@ def generate_lexers_json():
             file_hash = calculate_sha256(file_path)
             custom_lexers[filename] = file_hash
 
-    result = {"custom_lexers": custom_lexers}
+    return custom_lexers
 
-    with open(OUTPUT_JSON, "w") as file:
-        json.dump(result, file, indent=2)
+def generate_json_file():
+    # Load existing config, if it exists
+    config_path = Path(OUTPUT_JSON)
+    if config_path.exists():
+        with open(config_path, 'r') as f:
+            config = json.load(f)
+    else:
+        config = {}
+    
+    # Update only the lexers
+    config["custom_lexers"] = generate_lexers_json()
+
+    # Output the new config file
+    with open(config_path, "w") as file:
+        json.dump(config, file, indent=2)
 
     print(f"JSON file created at {OUTPUT_JSON}")
 
-
 if __name__ == "__main__":
-    generate_lexers_json()
+    generate_json_file()

### 2025-06-24 tools/generate_lexers_json.py
remove this directory's .latexminted_config in make clean and add security there

@@ -30,23 +30,27 @@ def generate_lexers_json():
 
     return custom_lexers
 
+
 def generate_json_file():
     # Load existing config, if it exists
     config_path = Path(OUTPUT_JSON)
     if config_path.exists():
-        with open(config_path, 'r') as f:
+        with open(config_path, "r") as f:
             config = json.load(f)
     else:
         config = {}
-    
+
     # Update only the lexers
     config["custom_lexers"] = generate_lexers_json()
+
+    config["security"] = {"enable_cwd_config": True}
 
     # Output the new config file
     with open(config_path, "w") as file:
         json.dump(config, file, indent=2)
 
     print(f"JSON file created at {OUTPUT_JSON}")
+
 
 if __name__ == "__main__":
     generate_json_file()

### 2025-06-24 tools/check_latexminted_config_exists.sh
nuke obsolete scripts

@@ -1,20 +1,0 @@
-#!/usr/bin/env bash
-
-FILE='.latexminted_config'
-
-if [[ ! ( -f "$HOME/$FILE" || -f "$TEXMF/$FILE" ) ]]; then
-    echo 'Error:'
-    echo "  Neither $HOME/$FILE (~/$FILE) nor $TEXMF/$FILE (\$TEXMF/$FILE) seem to exist!"
-    echo 
-    echo '  To create the file, please run: '
-    echo '    ./tools/create_latexminted_config.sh'  
-    echo ''
-    echo '  Or, manually create the file, with at least the following options set:'
-    echo
-    echo '  {'
-    echo '    "security": {' 
-    echo '      "enable_cwd_config": true' 
-    echo '    }' 
-    echo '  }' 
-    exit 1
-fi

### 2025-06-24 tools/create_latexminted_config.sh

@@ -1,13 +1,0 @@
-#!/usr/bin/env bash
-
-FILE="${HOME}/.latexminted_config"
-
-if [[ -f "$FILE" ]]; then
-    echo "Error: ${FILE} already exists"
-fi
-
-echo '{
-  "security": {
-      "enable_cwd_config": true
-  }
-}' > "$FILE"
